### 2016-05-04 paper/00specification.tex
specifications

\documentclass[11pt]{article}

\usepackage[backgroundcolor=white,textsize=small,linecolor=red]{todonotes}

\usepackage[margin=30mm]{geometry}
\usepackage{natbib}

%--------------------
% Math
\usepackage[fleqn]{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{multicol}
\usepackage{graphicx}

\newcommand{\mvalueof}[1]{\llbracket#1\rrbracket}
\newcommand{\citeposs}[2][]{\citeauthor{#2}'s (\citeyear[#1]{#2})}
\newcommand{\tuple}[1]{\ensuremath{\left\langle #1 \right\rangle}}

\newcommand{\hloranj}[1]{\textcolor[rgb]{.8,.33,.0}{#1}}% prints in orange
\newcommand{\argmax}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{max}}\;}
\newcommand{\argmin}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{min}}\;}
\newcommand{\sbna}{\exists\lnot\forall}

%--------------------
%\usepackage{gb4e}
%
%\usepackage{setspace}
%\onehalfspacing
%
%\usepackage{soul} %underlining
%\setuldepth{a}

%\usepackage{inconsolata}

%-------------------

\title{Bounds, Semantics/Pragmatics-Divide, and Evolution}

\author{%\bf NAME1 and NAME2\\
    \today
}

\date{}

\begin{document}
\maketitle

\section{Short exposition}
The plan is to expand our work on the semantics/pragmatics divide by means of a (more extended) study on the lack of upper-bounds in the literal meaning of scalar expressions. The main idea of this extension, other than providing a more detailed treatment as well as a comparison with iterated learning, is to look at the effect multiple scalar pairs have on the dynamics. That is, we consider larger language fragments consisting of more than one pair of expressions of the ``some but not all'' vs. ``all''-form.

\section{Differences to CogSci setup.}
\subparagraph{(I) Lexica and types.} $3$ pairs of scalar expressions. Each pair is a $(2,2)$-Boolean matrix, which means that a lexicon is now a $(8,8)$-matrix. We have two types of signaling behavior, literal and Gricean, and $6$ different $(2,2)$-lexica. The latter is the set of possible semantics of a single scalar pair. With three scalar pairs we have $6 \cdot 6 \cdot 6 = 216$ $(8,8)$-lexica. $216$ times two signaling behaviors makes for a total of $432$ types.

\subparagraph{(II) Observations.} Before, the set of all production observations was $O =$\linebreak  $\{\tuple{\tuple{s_1,m_i},\tuple{s_2,m_j}} | m_i, m_j \in M\}$. A member of $O$ encodes that a teacher produced $m_i$ in state $s_1$ and $m_j$ in $s_2$, i.e. it encodes one witnessed message for each state. A datum $d$ was a sequence of length $k$ of members of $O$. Learners witnessed such data sequences. Now, more in line with \citet{griffiths+kalish:2007}, $O = \{\tuple{s_i,m_j} | s_i \in S, m_j \in M\}$ and $d$ is a sequence of length $k$ of members of $O$. The main difference is that now some $d$ do not provide any production information for some states.

\subparagraph{(III) Computational load.} Computing the $Q$-matrix from the set of all observation sequences of length $k$ is computationally intractable for large $k$. Here, we follow \citet{griffiths+kalish:2007} as well in sampling $1000$ sequences $d$ from the set of all sequences when $k > 4$. We use this sampled subset to compute $Q$ instead of the full set of sequences\footnote{Thomas: Assuming this solves the problem. My computer can't handle the computation even with sampling but a better computer may. I'm still waiting on people from my project to address my request for computation time on our server.}

\subparagraph{(IV) Mutation/Learning matrix.} If the results warrant it, classical (unweighted) learning instead of communal (weighted) learning.\footnote{Thomas: We can save the results for both cases in one go and decide on this later.}

\paragraph{Further modifications we may want to keep in mind (but currently don't make use of).}
\subparagraph{(I) Cost for pragmatic reasoning.} At least in the CogSci setup the effect of adding cost to pragmatic reasoning is unsurprising: High cost for pragmatic signaling lowers the prevalence of pragmatic types. Lexica that semantically encode an upper-bound benefit the most from this. However, the cost needed to be substantial to make the pragmatic English-like lexicon stop being the incumbent type (particularly when learning is communal).

\subparagraph{(II) Negative learning bias.} Instead of penalizing complex semantics (semantic upper-bounds) one may consider penalizing simple semantics (no upper-bounds). This is useful as a sanity check but also yields unsurprising results in the CogSci setup: The more learners are biased against simple semantics, the more prevalent are lexica that semantically encode upper-bounds.

\subparagraph{(III) Inductive bias.} A second learning bias that codifies the idea that lexica should be uniform, i.e. be biased towards either lexicalizing an upper-bound for all weaker alternatives in a scalar pair or for none.

\subparagraph{(IV)  Uncertainty.} The other advantage of non-upper bounded semantics lies in being non-committal to the negation of stronger alternatives when the speaker is uncertain. Adding this to the model requires the most changes to our present setup and some additional assumptions about the cues available to players to discern the speaker's knowledge about the state she is in.

\section{General setup}
\subsection{Part 1: Signaling.}

We consider a population of players with two signaling strategies, literal and Gricean (level $0$ and $1$ below), each equipped with one of $216$ lexicons. This yields a total of $432$ distinct player types $t \in T$. However, each lexicon is composed of three independent lexical pairs. That is, while $|M| = |S| = 8$, we can decompose each lexicon into three independent $(2,2)$-matrices. A lexicon is a combination of three such matrices. These are listed in Table \ref{tab:lexica}.

\begin{table}[h]
\centering 
\begin{tabular}{l c l}
$L_1$ = $\begin{pmatrix} 0 & 0 \\ 1 & 1 \end{pmatrix}$ & 
$L_2$ = $\begin{pmatrix} 1 & 1 \\ 0 & 0 \end{pmatrix}$ & 
$L_3$ = $\begin{pmatrix} 1 & 1 \\ 1 & 1 \end{pmatrix}$\\[0.5cm]

$L_4$ = $\begin{pmatrix} 0 & 1 \\ 1 & 0 \end{pmatrix}$ &
$L_5$ = $\begin{pmatrix} 0 & 1 \\ 1 & 1 \end{pmatrix}$ &
$L_6$ = $\begin{pmatrix} 1 & 1 \\ 1 & 0 \end{pmatrix}$
\end{tabular}
\caption{{\footnotesize Set of possible $(2,2)$-matrices, i.e. possible semantics for a pair of scalar expressions.}}
\label{tab:lexica}
\end{table}

As in the CogSci paper, $L_4$ (semantic upper-bound for $m_2$) and $L_5$ (no semantic upper-bound for $m_2$) are the target lexica. Gricean $L_5$ users can convey/infer the bound pragmatically, while literal/Gricean $L_4$ users do so semantically.

\paragraph{Signaling behavior.} With $\lambda \geq 1$ (rationality parameter), $\alpha \in [0,1]$ (pragmatic violations) and $pr \in \Delta(S)$ a common prior over $S$ (uniform so far):

\vspace{-0.15cm}
\begin{flalign}
&R_{0}(s|m;L) \propto pr(s) L_{sm}\label{litl}\\
&S_{0}(m|s;L) \propto \exp(\lambda \; L_{sm}) \label{lits}\\
&R_{1}(s|m;L) \propto pr(s) S_{0}(m|s;L) \label{pragl}\\
&S_{1}(m|s;L) \propto  \exp(\lambda \; R_{0}(s|m;L)^\alpha) \label{prags}
\end{flalign}

\paragraph{Symmetrized expected utility.} With $P \in \Delta(S)$ (uniform so far; therefore $P = pr$):
\begin{itemize}

  \item $U(t_i,t_j) = [U_S(t_i,t_j) + U_R(t_i,t_j)] / 2$
  \item $U_S(t_i,t_j) = \sum_s P(s) \; \sum_m P_S(m|s;t_i) \; \sum_{s'} P_R(s'|m,t_j) \; \delta(s,s')$, where $\delta(s,s')$ returns $1$ iff $s = s'$ and otherwise $0$
  \item $U_R(t_i,t_j) = U_S(t_j,t_i)$
\end{itemize}

\subsection{Part 2: Mutation \& replication.}
The proportion of players of type $i$, $x_i$, is initialized as an arbitrary distribution over $T$. $p^\star \in \Delta(T)$ is learning a prior over (player) types dependent only on the lexicon of the type. 
\begin{itemize}
    \item $f_i = \sum_j x_j U(x_i,x_j)$
    \item $\Phi = \sum_i x_i f_i$
    \item $Q_{ij} \propto \sum_d P(d|t_i) \; P(t_j|d)$, where $P(t_j|d) \propto p^\star(t_j) P(d|t_j)$ and $d$ is a sequence of observations of length $k$ of the form \tuple{\tuple{s_i,m_j}, ... \tuple{s_k, m_l}}.
\end{itemize}

\paragraph{Weighted (communal) learning vs. unweighted (parental) learning}

\begin{itemize}
	\item For parental learning (standard RMD): $\dot x_i = \sum_j Q_{ji} \frac{x_j f_j}{\Phi}$
	\item For communal learning (our variant in CogSci): $\dot x_i = \sum_j N_{ji} \frac{x_j f_j}{\Phi}$, where $N_{ji} = x'_{i} Q_{ji}$ and $x'_i$ is the proportion of type $i$ in the previous generation.
\end{itemize}

\subsection{Model parameters \& procedure} 
\begin{enumerate}
  \item Sequence length $k$
  \item Pragmatic production parameter $\alpha$
  \item Rationality parameter $\lambda$
  \item Learning prior over types (lexica); cost parameter $c$. $p^\star(t_i) \propto n - c \cdot r$ where $n$ is the total number of states and $r$ that of upper-bounded messages only true of $s_1$ in $t_i$'s lexicon (if only $s_1$ is true of a message, then this message encodes an upper-bound). Then the score for $L_1$, $L_3$, $L_5$ is $2$, that of $L_4$ and $L_6$ is $2-c$, and that of $L_2$ is $2-2c$; Normalization over lexica scores yields the prior over lexica (which is equal to the prior over types).   
  \item Prior over meanings ($pr$). We assume that $pr(s) = \frac{1}{|S|}$ for all $s$.
  \item True state distribution ($P$). We currently assume that $P = \frac{1}{|S|}$ but it may be interesting to vary this
\end{enumerate}

\paragraph{Procedural description.} The game is initialized with some arbitrary distribution over player types. At the game's onset we compute $Q$ once based on the set of sequences $D$. If $k > 4$, $1000$ sequences are sampled from $D$ and used to compute $Q$ (instead of the full set). Replicator dynamics are computed based on the fitness of each type in the current population as usual. In the case of community learning, $Q$ is weighted by the previous population vector in each new generation.

\paragraph{Presentation.} We have a large number of types, some of which are highly similar. For instance, a type that uses $L_4$, $L_5$ and $L_4$ for her respective three scalar pairs is a different type than one that uses $L_5$, $L_4$ and $L_4$. As our goal is to elucidate the role and prevalence of upper-bounds, we may summarize the results not by type but by lexica counts. For instance, if the previous two types where all we had this would mean that $L_4$'s proportion is roughly $.66$ and $L_5$'s $.33$ (irrespective of signaling behavior). An example is given below for communal learning with a single scalar pair against three scalar pairs ($\alpha = 1, c = .7, \lambda = 30, k = 3$):

\begin{figure}[h] %htb for exact position
	\centering
	\resizebox{13cm}{!}{\includegraphics{../plots/wgh-single-vs-multi-k3-c70}}
\end{figure}

%\bibliographystyle{apacite}
\bibliographystyle{unsrtnat}

%\setlength{\bibleftmargin}{.125in}
%\setlength{\bibindent}{-\bibleftmargin}
\bibliography{./bounds-rmd}

\end{document}

### 2016-06-05 IL-comparison/IL-comparison.tex
comparison with IL*

\documentclass[12pt]{article}

\usepackage{graphicx}

\usepackage[round]{natbib}
\usepackage{hyperref}
\usepackage{color}
\usepackage{gb4e}
\usepackage{subcaption}
\usepackage{multicol}

\usepackage{stmaryrd}

\newcommand{\mvalueof}[1]{\llbracket#1\rrbracket}

%%%New commands%%%
\newcommand{\tuple}[1]{\ensuremath{\left\langle #1 \right\rangle}} 
\newcommand{\citeposs}[2][]{\citeauthor{#2}'s (\citeyear[#1]{#2})}
\newcommand{\hloranj}[1]{\textcolor[rgb]{.8,.33,.0}{#1}}% prints in orange
\newcommand{\hlblue}[1]{\textcolor[rgb]{0.13,0.67,0.8}{#1}}
\newcommand{\hlgrey}[1]{\textcolor[rgb]{0.57, 0.64, 0.69}{#1}}

%\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\argmax}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{max}}\;}
\newcommand{\argmin}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{min}}\;}

\newcommand*{\defeq}{\mathrel{\vcenter{\baselineskip0.5ex \lineskiplimit0pt
                     \hbox{\scriptsize.}\hbox{\scriptsize.}}}%
                     =}

	%%%%%%%%%%%%%%%%%%
\usepackage{mathcomp}
\usepackage{blkarray} %to label matrix
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{nicefrac}

\begin{document}

\title{(Evolutionary) Game Theory meets Bayesian (Iterated) Learning}
\date{\today}

\maketitle
\tableofcontents

\abstract{Bayesian Game Theory as a framework to analyze language change}

\section{Introduction}
Natural languages change over time both within and across generations. That is, this process is influenced both by synchronic and diachronic developments. A number of different proposals have been put forward to explain the emergence and persistence of core features of language, such as \hloranj{add references e.g. for compositionality.} Notwithstanding, while differences can be gleamed from particular analyzes, the overall argument is one of competing pressures. On the one hand, languages need to be expressive enough to ensure communicative success within a linguistic community. On the other hand, they should be well-adapted to their faithful transmission across generations. In other words, functional pressure in tandem with learnability are expected to shed light on properties that allow natural language to cope with these forces.

Evolutionary game theory is one of the frameworks that has been used to analyze the dynamics of such competing forces \citep{nowak+krakauer:1999,huttegger+zollman:2013}. In the following, we focus on the replicator-mutator dynamics (RMD) which \hloranj{short blurb on this} (for an overview see \citealt{hofbauer+sigmund:2003}. Under this view, developments in the cultural evolution of languages are investigated in analogy to biological forces of evolution such as selection and mutation. In particular, the fitness of a geno- or phenotype is cast in terms of the communicative success of a type of communicative behavior, a core of which is the language spoken by it. The main assumption being that communicatively well-adapted languages prosper and spread in a population, or, framed in biological terms, produce more offspring. Mutation, in turn, is viewed as the faithful transmission of a language or communicative behavior, rather than the faithful transmission of biological information, i.e. its acquisition by a new generation of speakers.

A different view is offered by the iterated learning paradigm \hloranj{citations}. While the overall focus is still one of competing pressures, the focus lies in the explanatory value of language acquisition. In these models, languages are transmitted through learner-speaker chains. At the beginning of the chain, the first learner is exposed to some linguistic data from which she infers a language and produces linguistic data which then serves as input for the following learner in the chain. Iterated learning has been argued to yield insights in \hloranj{phenomena and citations}. One of its advantages is its versatility, being used for analytic analyzes, numerical simulations and human experiments \hloranj{more citations}. However, ... summarize \hloranj{Griffiths+kalish}.

In the following, \hloranj{give an overview}.

\section{Components of RMD}

\subsection{Interpretation of components}

The linguistic interpretation of the replicator-mutator dynamics and evolutionary game theory more broadly has stayed, by en large, close to its biological counterpart. The advantage of this broad view lies in the generality of the results obtained and enables for a straightforward import of insights gained from biological evolutionary game theory to its linguistic application, and vice versa. However, we believe that a more language-centric interpretation gives us additional leverage to understand, compare, and modify our models. On the technical side our proposal is conservative in that the dynamics of RMD remain as before. The difference from past approaches lies in the representation of mutation as a process of iterated Bayesian learning, and the final generation being analyzed after a replication instead of a mutation phase. We hope that the incorporation of (iterated) Bayesian learning in evolutionary models will allow to bridge and synthesize the results from different areas of inquiry on the cultural evolution of language. In particular, on the one side, language game paradigms as \citeposs{lewis:1969} signaling games and \citeposs{steels:2012} naming games, and on the other, iterated learning \citep{kirby:2001}.

Classically, replication is framed in terms of reproduction (based on communicative success) and mutation as language acquisition error (see e.g. \citealt{komarova+etal:2001,nowak+etal:2001, nowak+etal:2002,hutteger+zollman:2013}). The former indicates the proportion of offspring a type of language user produces. The better adapted for communication within a population a type is, the more children it produces. The latter represents the proportion of children of a type that will faithfully adopt the type of their parents -- or adopt a different type. This introduces perturbations in the transmission of a language across generations and controls the proportion of transitions from one type to another. Mutation is often qualified as noise or confusability, and can either be uniform or dependent on any two types, i.e. either as a fixed chance of not learning a parent's language, adopting another uniformly, or be determined by some metric of closeness between two languages whereby close languages are more likely to be confused by learners.\footnote{\hloranj{They do generally compare different learning mechanisms. Either batch or sequential learning. Double check where learning comes in.}}

\hloranj{Continue with fitness as synchronic adaptation. Fitness as condenced summary of adaption in synchronic change}

\hloranj{Biological transmission, framed in terms of cultural transmission is naturally understood as learning}

\paragraph{Fitness} Summary of adoption in synchronic change. \hloranj{cf. Komarova et al. 2001, Nowak et al 2001, 2002 for possible interpretations of fitness in evo ling settings}
\paragraph{Mutation} Learning

\subparagraph{Prior} Amount of evidence
\subparagraph{Data} Difficult. Community? Not? Parent? All observations?
\subparagraph{Learning} Matching? Maximum?

\section{Comparisons}
\subsection{Comparison to IL}

Main result: Stationary distribution of IL is not reached by BGT.
\subsection{MAP and RD without M}
Does MAP turn into RD? No, because input is still important.

\section{Toy example}
For comparability, we follow the setup analyzed by \citet[\S 6]{griffths+kalish:2007}. $S = \{00,01,10,11\} = M$. $\mathcal{L}$ is the set of all possible mappings from $S$ to $M$, i.e. $4^4 = 256$ possible signaling systems.\footnote{Note that \citet{griffiths+kalish:2007} only take one-to-one mappings into consideration.} Four of which are compositional in that meaning and form components agree position-wise.\footnote{Note that \citet{griffiths+kalish:2007} take $260$ languages into consideration. $256$ possible (holistic) languages and $4$ compositional ones. However, the former set includes $4$ `holistic' languages identical to the $4$ compositional ones. We see no principled reason to consider such identical languages as distinct. As a consequence, we consider $252$ holistic and $4$ compositional languages.} For simplicity syntax, i.e. component order, is assumed to be given. $P(s) = \nicefrac{1}{|S|}$. Prior parameter $\alpha \in [.01,0.5]$, error parameter $\epsilon \in [.01,.05]$ and data length $n \in [1,10]$.

\begin{align}
P(m|s,L) &= \left\{\begin{array}{l} 1 - \epsilon \qquad \textnormal{if $m$ is true of $s$ in $L$}\\
\frac{\epsilon}{3} \qquad \textnormal{otherwise}\end{array}\right.
%U_S(t_i, m_j, a_k) &= U_R(t_i, m_j, a_k)
\end{align}

Hierarchical prior:
\begin{align}
P(L) &= \left\{\begin{array}{l} \frac{\alpha}{4} \qquad \textnormal{if $L$ is compositional}\\
\frac{1 - \alpha}{252} \qquad \textnormal{otherwise}\end{array}\right.
%U_S(t_i, m_j, a_k) &= U_R(t_i, m_j, a_k)
\end{align}

Production for EU:
\begin{align}
P(s|m,L) &= \left\{\begin{array}{l} \frac{1}{|\{s | \textnormal{$m$ is true of $s$ in $L$}\}|} \qquad \textnormal{if $m$ is true of $s$ in $L$}\\
\frac{\epsilon}{|\{s | \textnormal{$m$ is false of $s$ in $L$}\}|} \qquad \textnormal{otherwise}\end{array}\right.
%U_S(t_i, m_j, a_k) &= U_R(t_i, m_j, a_k)
\end{align}

\paragraph{Results:} Compute probability matching and MAP for IL and RMD with iterated Bayesian learning. Save both last mutation and last replication as results.

\section{Possibly add synchronic interaction}

\section{Conclusion}

\bibliographystyle{unsrtnat}

\bibliography{../../../../../../tex/masternolinks}

\end{document}

### 2016-06-05 IL-comparison/Player.py

@@ -1,0 +1,32 @@
+import operator
+import math, random
+import numpy as np
+import copy
+#import sys
+#            sys.exit('error')
+
+def normalize(m):
+    m = m / m.sum(axis=1)[:, np.newaxis]
+    m[np.isnan(m)] = 0.
+    return m
+
+
+class Player:
+    def __init__(self,lexicon,epsilon):
+        self.states = [00,01,10,11]
+        self.messages = [00,01,10,11]
+        self.lexicon = lexicon
+        self.epsilon = epsilon
+
+    def senderMatrix(self):
+        out = np.zeros((len(self.states),len(self.messages)))
+        for i in range(np.shape(out)[0]):
+            for j in range(np.shape(out)[1]):
+                if self.lexicon[i,j] == 1:
+                    out[i,j] = 1 - self.epsilon
+                else:
+                    out[i,j] = self.epsilon / (len(self.messages) - 1)
+        return out
+
+    def receiverMatrix(self):
+        return normalize(np.transpose(self.senderMatrix()))

### 2016-06-05 IL-comparison/plot.py

@@ -1,0 +1,224 @@
+###Plots##
+#Reordering the array so that compositional languages are at indices 0-3:
+c3a = [results3a[compIndices[0]],results3a[compIndices[1]], results3a[compIndices[2]],results3a[compIndices[3]]]
+c3b = [results3b[compIndices[0]],results3b[compIndices[1]], results3b[compIndices[2]],results3b[compIndices[3]]]
+c3c = [results3c[compIndices[0]],results3c[compIndices[1]], results3c[compIndices[2]],results3c[compIndices[3]]]
+
+for i in range(len(compIndices)): 
+    results3a = np.delete(results3a,compIndices[i])
+    results3a = np.insert(results3a,i,c3a[i])
+
+    results3b = np.delete(results3b,compIndices[i])
+    results3b = np.insert(results3b,i,c3b[i])
+
+    results3c = np.delete(results3c,compIndices[i])
+    results3c = np.insert(results3c,i,c3c[i])
+    
+###Adding some white space to separate C from H:
+for i in range(len(compIndices),len(compIndices)+10):
+    results3a = np.insert(results3a,i,0)
+    results3b = np.insert(results3b,i,0)
+    results3c = np.insert(results3c,i,0)
+
+
+X = np.arange(len(hypotheses)+10)
+Ya = np.sqrt(results3a)
+Yb = np.sqrt(results3b)
+Yc = np.sqrt(results3c)
+
+ymin,ymax = 0,np.max(np.array([Ya[np.argmax(results3a)],Yb[np.argmax(results3b)],Yc[np.argmax(results3c)]]))+0.01
+
+fig, ax = plt.subplots()
+ax.set_xticks(np.array([len(compIndices)/2,(len(hypotheses)-len(compIndices))/2]))
+ax.set_xticklabels(('compositional', 'holistic'))
+plt.ylabel('Mean in population')
+ax.bar(X,Ya,width=1)
+ax.margins(0.025, 0.025)
+ax.text(.5,.9,'Functional pressure and iterated learning',
+        horizontalalignment='center',
+        transform=ax.transAxes)
+ax.set_ylim([ymin,ymax])
+plt.show()
+
+fig, ax = plt.subplots()
+ax.set_xticks(np.array([len(compIndices)/2,(len(hypotheses)-len(compIndices))/2]))
+ax.set_xticklabels(('compositional', 'holistic'))
+ax.bar(X,Yb)
[…]

### 2016-06-05 IL-comparison/rmd-bayes.py

@@ -1,0 +1,281 @@
+import numpy as np
+#np.set_printoptions(threshold=np.nan)
+from itertools import product,permutations
+from random import choice
+from Player import Player
+from math import exp,sqrt
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set_context(rc={'lines.markeredgewidth': 0.5})
+import sys 
+from copy import copy
+from collections import defaultdict
+import datetime
+import csv
+
+def normalize(m):
+    return m / m.sum(axis=1)[:, np.newaxis]
+
+def norm(m):
+    m = m / m.sum(axis=1)[:, np.newaxis]
+    m[np.isnan(m)] = 1. / np.shape(m)[1]
+    return m
+
+#lexicon = np.array([[1,0,1],[0,1,1]])
+s = 4 #number of states
+n = 3 #observation sequence length
+generations = 100 #generations per game
+games = 5 #amount of independent games
+alpha = 0.01 #bias parameter
+epsilon = 0.05
+mc = 1000 #amount of samples
+
+
+f_il_samp_mean = csv.writer(open('./results/comparison-il-samp-mean-a%d-e%f-k%d-g%d-r%d.csv' %(alpha,epsilon,n,generations,games),'wb')) 
+f_unwgh_samp_mean = csv.writer(open('./results/comparison-unwgh-samp-mean-a%d-e%f-k%d-g%d-r%d.csv' %(alpha,epsilon,n,generations,games),'wb')) 
+f_wgh_samp_mean = csv.writer(open('./results/comparison-wgh-samp-mean-a%d-e%f-k%d-g%d-r%d.csv' %(alpha,epsilon,n,generations,games),'wb')) 
+
+f_il_map_mean= csv.writer(open('./results/comparison-il-samp-map-a%d-e%f-k%d-g%d-r%d.csv' %(alpha,epsilon,n,generations,games),'wb')) 
+f_unwgh_map_mean= csv.writer(open('./results/comparison-unwgh-map-mean-a%d-e%f-k%d-g%d-r%d.csv' %(alpha,epsilon,n,generations,games),'wb')) 
+f_wgh_map_mean = csv.writer(open('./results/comparison-wgh-map-mean-a%d-e%f-k%d-g%d-r%d.csv' %(alpha,epsilon,n,generations,games),'wb')) 
+
[…]

### 2016-06-15 IL-comparison/IL-comparison.tex
learning observations by production sampling

@@
 
-\title{(Evolutionary) Game Theory meets Bayesian (Iterated) Learning}
+\title{Comparison between iterated learning and fitness+learning}
 \date{\today}
@@
 \maketitle
-\tableofcontents
 
-\abstract{Bayesian Game Theory as a framework to analyze language change}
-
-\section{Introduction}
-Natural languages change over time both within and across generations.
-That is, this process is influenced both by synchronic and diachronic developments.
-A number of different proposals have been put forward to explain the emergence and persistence of core features of language, such as \hloranj{add references e.g.
-for compositionality.} Notwithstanding, while differences can be gleamed from particular analyzes, the overall argument is one of competing pressures.
-On the one hand, languages need to be expressive enough to ensure communicative success within a linguistic community.
-On the other hand, they should be well-adapted to their faithful transmission across generations.
-In other words, functional pressure in tandem with learnability are expected to shed light on properties that allow natural language to cope with these forces.
-
-Evolutionary game theory is one of the frameworks that has been used to analyze the dynamics of such competing forces \citep{nowak+krakauer:1999,huttegger+zollman:2013}.
-In the following, we focus on the replicator-mutator dynamics (RMD) which \hloranj{short blurb on this} (for an overview see \citealt{hofbauer+sigmund:2003}.
-Under this view, developments in the cultural evolution of languages are investigated in analogy to biological forces of evolution such as selection and mutation.
-In particular, the fitness of a geno- or phenotype is cast in terms of the communicative success of a type of communicative behavior, a core of which is the language spoken by it.
-The main assumption being that communicatively well-adapted languages prosper and spread in a population, or, framed in biological terms, produce more offspring.
-Mutation, in turn, is viewed as the faithful transmission of a language or communicative behavior, rather than the faithful transmission of biological information, i.e.
-its acquisition by a new generation of speakers.
-
-A different view is offered by the iterated learning paradigm \hloranj{citations}.
-While the overall focus is still one of competing pressures, the focus lies in the explanatory value of language acquisition.
-In these models, languages are transmitted through learner-speaker chains.
-At the beginning of the chain, the first learner is exposed to some linguistic data from which she infers a language and produces linguistic data which then serves as input for the following learner in the chain.
-Iterated learning has been argued to yield insights in \hloranj{phenomena and citations}.
-One of its advantages is its versatility, being used for analytic analyzes, numerical simulations and human experiments \hloranj{more citations}.
-However, ...
-summarize \hloranj{Griffiths+kalish}.
-
-In the following, \hloranj{give an overview}.
-
-\section{Components of RMD}
-
-\subsection{Interpretation of components}
-
-The linguistic interpretation of the replicator-mutator dynamics and evolutionary game theory more broadly has stayed, by en large, close to its biological counterpart.
-The advantage of this broad view lies in the generality of the results obtained and enables for a straightforward import of insights gained from biological evolutionary game theory to its linguistic application, and vice versa.
-However, we believe that a more language-centric interpretation gives us additional leverage to understand, compare, and modify our models.
-On the technical side our proposal is conservative in that the dynamics of RMD remain as before.
-The difference from past approaches lies in the representation of mutation as a process of iterated Bayesian learning, and the final generation being analyzed after a replication instead of a mutation phase.
-We hope that the incorporation of (iterated) Bayesian learning in evolutionary models will allow to bridge and synthesize the results from different areas of inquiry on the cultural evolution of language.
-In particular, on the one side, language game paradigms as \citeposs{lewis:1969} signaling games and \citeposs{steels:2012} naming games, and on the other, iterated learning \citep{kirby:2001}.
-
-Classically, replication is framed in terms of reproduction (based on communicative success) and mutation as language acquisition error (see e.g.
-\citealt{komarova+etal:2001,nowak+etal:2001, nowak+etal:2002,hutteger+zollman:2013}).
-The former indicates the proportion of offspring a type of language user produces.
-The better adapted for communication within a population a type is, the more children it produces.
-The latter represents the proportion of children of a type that will faithfully adopt the type of their parents -- or adopt a different type.
-This introduces perturbations in the transmission of a language across generations and controls the proportion of transitions from one type to another.
-Mutation is often qualified as noise or confusability, and can either be uniform or dependent on any two types, i.e.
-either as a fixed chance of not learning a parent's language, adopting another uniformly, or be determined by some metric of closeness between two languages whereby close languages are more likely to be confused by learners.\footnote{\hloranj{They do generally compare different learning mechanisms.
-Either batch or sequential learning.
-Double check where learning comes in.}}
-
-\hloranj{Continue with fitness as synchronic adaptation. Fitness as condenced summary of adaption in synchronic change}
-
-\hloranj{Biological transmission, framed in terms of cultural transmission is naturally understood as learning}
-
-\paragraph{Fitness} Summary of adoption in synchronic change. \hloranj{cf. Komarova et al. 2001, Nowak et al 2001, 2002 for possible interpretations of fitness in evo ling settings}
-\paragraph{Mutation} Learning
-
-\subparagraph{Prior} Amount of evidence
-\subparagraph{Data} Difficult. Community? Not? Parent? All observations?
-\subparagraph{Learning} Matching? Maximum?
-
-\section{Comparisons}
-\subsection{Comparison to IL}
-
-Main result: Stationary distribution of IL is not reached by BGT.
-\subsection{MAP and RD without M}
-Does MAP turn into RD? No, because input is still important.
-
-\section{Toy example}
-For comparability, we follow the setup analyzed by \citet[\S 6]{griffths+kalish:2007}.
+For comparability, we follow the setup analyzed by \citet[\S 6]{griffiths+kalish:2007} on the emergence of compositionality.
 $S = \{00,01,10,11\} = M$.
@@
 However, the former set includes $4$ `holistic' languages identical to the $4$ compositional ones.
-We see no principled reason to consider such identical languages as distinct.
-As a consequence, we consider $252$ holistic and $4$ compositional languages.} For simplicity syntax, i.e.
+In contrast, we consider $252$ holistic and $4$ compositional languages.
+(N.B.: They include 4 languages holistic languages indistinguishable from their compositional counterparts to show that the learning bias drives selection.
+We already know that, so it's not necessary.)} Syntax, i.e.
 component order, is assumed to be given.
 $P(s) = \nicefrac{1}{|S|}$.
-Prior parameter $\alpha \in [.01,0.5]$, error parameter $\epsilon \in [.01,.05]$ and data length $n \in [1,10]$.
-
-\begin{align}
-P(m|s,L) &= \left\{\begin{array}{l} 1 - \epsilon \qquad \textnormal{if $m$ is true of $s$ in $L$}\\
-\frac{\epsilon}{3} \qquad \textnormal{otherwise}\end{array}\right.
-%U_S(t_i, m_j, a_k) &= U_R(t_i, m_j, a_k)
-\end{align}
+The simulations involve a (hierarchical) prior parameter $\alpha \in [.01,0.5]$, an error parameter $\epsilon \in [.01,.05]$ and data length $n \in [1,10]$.
+\vspace{1cm}
 
@@
 P(L) &= \left\{\begin{array}{l} \frac{\alpha}{4} \qquad \textnormal{if $L$ is compositional}\\
-\frac{1 - \alpha}{252} \qquad \textnormal{otherwise}\end{array}\right.
+\frac{1 - \alpha}{252} \quad \textnormal{otherwise}\end{array}\right.
 %U_S(t_i, m_j, a_k) &= U_R(t_i, m_j, a_k)
@@
 
-Production for EU:
+Production: 
 \begin{align}
-P(s|m,L) &= \left\{\begin{array}{l} \frac{1}{|\{s | \textnormal{$m$ is true of $s$ in $L$}\}|} \qquad \textnormal{if $m$ is true of $s$ in $L$}\\
-\frac{\epsilon}{|\{s | \textnormal{$m$ is false of $s$ in $L$}\}|} \qquad \textnormal{otherwise}\end{array}\right.
+P(m|s,L) &= \left\{\begin{array}{l} 1 - \epsilon \qquad \textnormal{if $m$ is true of $s$ in $L$}\\
+\frac{\epsilon}{3} \quad \  \qquad \textnormal{otherwise}\end{array}\right.
 %U_S(t_i, m_j, a_k) &= U_R(t_i, m_j, a_k)
@@
 
-\paragraph{Results:} Compute probability matching and MAP for IL and RMD with iterated Bayesian learning. Save both last mutation and last replication as results.
+Interpretation (not covered by \citealt{griffiths+kalish:2007}):
+\begin{align}
+P(s|m,L) \propto P(s) P(m|s,L)
+\end{align}
 
-\section{Possibly add synchronic interaction}
+\paragraph{Results.} The following plots exemplify the main differences between iterated learning and evolutionary game theory with Bayesian learning.
+Each dynamic was run with learning as sampling from the posterior ({\em post-sampling}) as well as with learning as maximum posterior likelihood ({\em MAP}).
+Furthermore, we include results for both parental learning (standard mutation) and communal learning (weighted mutation) for the EGT variant.
+All quantities were subjected to a square root transformation to make the differences between these variants more apparent-- in particular for the quantities of holistic languages that remain under each variant.
+The parameter values correspond to those reported by \citealt[466]{griffiths+kalish:2007} (rightmost plots), therefore, the results from iterated learning with posterior sampling as well as the prior double as a replication of their setup.
 
-\section{Conclusion}
+To ensure comparability, we also follow Griffiths \& Kalish in running all simulations for $10000$ generations and taking an average of $1000$ such games for each parameter value and variant of dynamics/learning.
+
+\paragraph{Simulation 1.} $\alpha = 0.5, \epsilon = 0.05, n = 1$. Observations:
+
+\begin{figure}[h!]
+\makebox[\linewidth][c]{%
+\begin{subfigure}[b]{.7\textwidth}
+\centering
+\includegraphics[scale=0.5]{./plots/test1}
+%\caption{a test subfigure}
+\end{subfigure}%
+\begin{subfigure}[b]{.7\textwidth}
+\centering
+\includegraphics[scale=0.5]{./plots/test2}
+%\caption{a test subfigure}
+\end{subfigure}%
+}
+\end{figure}
+
+\paragraph{Simulation 2.} $\alpha = 0.5, \epsilon = 0.05, n = 1$
+\paragraph{Simulation 3.} $\alpha = 0.5, \epsilon = 0.05, n = 1$
+\paragraph{Simulation 4.} $\alpha = 0.5, \epsilon = 0.05, n = 1$
 
@@
 
-\bibliography{../../../../../../tex/masternolinks}
+\bibliography{../../Dropbox/tex/masternolinks}
 

### 2016-06-15 1plot.py

@@ -8,8 +8,8 @@ sns.set_context(rc={'lines.markeredgewidth': 0.5})
 import csv
 import sys
 
-f1 = csv.reader(open('./results/singlescalar-unwgh-mean-a1-c0.900000-l30-k14-g30-r100.csv','rt'))
-f2 = csv.reader(open('./results/1multiscalar-unwgh-mean-a1-c0.900000-l30-k14-g30-r100.csv','rt'))
+f1 = csv.reader(open('./results/singlescalar-wgh-mean-a1-c0.100000-l30-k9-g30-r100.csv','rt'))
+f2 = csv.reader(open('./results/1multiscalar-wgh-mean-a1-c0.100000-l30-k9-g30-r100.csv','rt'))
 
 def results_by_lexica_multi(f):
     p = np.zeros(6)
@@ -61,7 +61,7 @@ ax.margins(0.025, 0.025)
 #ax.text(.5,.9,'$\alpha$ = 1, c = %r, $\lambda$ = %r, pairs = %r, k = %r',
 #        horizontalalignment='center',
 #        transform=ax.transAxes) % (alpha, cost, lam, lexical_pairs, k)
-ax.set_title('Unweighted mutation. a = %r, c = %r, lam = %r, k = %r' % (1, .9, 30, 14))
+ax.set_title('Weighted mutation. a = %r, c = %r, lam = %r, k = %r' % (1, .1, 30, 9))
 
 #ax.set_ylim([ymin,ymax])
 plt.show()

### 2016-06-15 4-parental-separateobs-rmd-multiscalar.py

@@ -4,12 +4,15 @@
 #2 possible signaling behaviors: literal or gricean
 #12 types (6 lexica * 2 signaling behaviors) per signaling pair
 #6*6*6 = 216  literal types, 216 gricean types, 432 types in total
+#Set of observations of length k:
+##Sample <s,m> tuples from a type's production probabilities, P(a = <s,m> | t_i) = P(s) * P(m|s,t_i)
+#
 #####
 
 
 import numpy as np
 #np.set_printoptions(threshold=np.nan)
-from random import sample
+from random import sample,choice
 from itertools import product, permutations, combinations_with_replacement
 from player import LiteralPlayer,GriceanPlayer
 import sys 
@@ -18,24 +21,30 @@ import csv
 
 
 alpha = 1 # rate to control difference between semantic and pragmatic violations
-cost = 0.9 # cost for LOT-concept with upper bound
+cost = 0.4 # cost for LOT-concept with upper bound
 lam = 30 # soft-max parameter
-k = 3  # number of learning observations
-mc = 1000 #amount of samples from OBS when k > 4
+k = 10  # observation sequence length 
+sample_amount = 100 #amount of samples per type production
 
-gens = 30 #number of generations per simulation run
+gens = 50 #number of generations per simulation run
 runs = 100 #number of independent simulation runs
-
 
 lexical_pairs = 3 #number of 2x2 matrices
 states = 2 #number of states
 messages = 2 #number of messages
 
+learning_parameter = 5 #prob-matching = 1, increments approach MAP
 
-f_unwgh_mean = csv.writer(open('./results/multiscalar-unwgh-mean-a%d-c%f-l%d-k%d-g%d-r%d.csv' %(alpha,cost,lam,k,gens,runs),'wb')) #file to store each unweighted simulation run after n generations
-f_wgh_mean = csv.writer(open('./results/multiscalar-wgh-mean-a%d-c%f-l%d-k%d-g%d-r%d.csv' %(alpha,cost,lam,k,gens,runs),'wb')) #file to store each weighted simulation run after n generations
+state_freq = np.ones(states*lexical_pairs) / float(states*lexical_pairs) #frequency of states. 
[…]

### 2016-06-15 IL-comparison/Player.py

@@ -1,9 +1,4 @@
-import operator
-import math, random
 import numpy as np
-import copy
-#import sys
-#            sys.exit('error')
 
 def normalize(m):
     m = m / m.sum(axis=1)[:, np.newaxis]

### 2016-06-15 IL-comparison/il-comparison-rmd.py

@@ -1,15 +1,9 @@
 import numpy as np
 #np.set_printoptions(threshold=np.nan)
 from itertools import product,permutations
-from random import choice
-from Player import Player
-from math import exp,sqrt
-import matplotlib.pyplot as plt
-import seaborn as sns
-sns.set_context(rc={'lines.markeredgewidth': 0.5})
+from random import choice,sample
+from player import Player
 import sys 
-from copy import copy
-from collections import defaultdict
 import datetime
 import csv
 
@@ -23,10 +17,10 @@ def norm(m):
 
 #lexicon = np.array([[1,0,1],[0,1,1]])
 s = 4 #number of states
-n = 3 #observation sequence length
-generations = 100 #generations per game
-games = 5 #amount of independent games
-alpha = 0.01 #bias parameter
+n = 10 #observation sequence length
+generations = 10000 #generations per game
+games = 1000 #amount of independent games
+alpha = 0.5 #bias parameter
 epsilon = 0.05
 mc = 1000 #amount of samples
 
@@ -35,7 +29,7 @@ f_il_samp_mean = csv.writer(open('./results/comparison-il-samp-mean-a%d-e%f-k%d-
 f_unwgh_samp_mean = csv.writer(open('./results/comparison-unwgh-samp-mean-a%d-e%f-k%d-g%d-r%d.csv' %(alpha,epsilon,n,generations,games),'wb')) 
 f_wgh_samp_mean = csv.writer(open('./results/comparison-wgh-samp-mean-a%d-e%f-k%d-g%d-r%d.csv' %(alpha,epsilon,n,generations,games),'wb')) 
 
-f_il_map_mean= csv.writer(open('./results/comparison-il-samp-map-a%d-e%f-k%d-g%d-r%d.csv' %(alpha,epsilon,n,generations,games),'wb')) 
+f_il_map_mean= csv.writer(open('./results/comparison-il-map-mean-a%d-e%f-k%d-g%d-r%d.csv' %(alpha,epsilon,n,generations,games),'wb')) 
 f_unwgh_map_mean= csv.writer(open('./results/comparison-unwgh-map-mean-a%d-e%f-k%d-g%d-r%d.csv' %(alpha,epsilon,n,generations,games),'wb')) 
[…]

### 2016-06-15 IL-comparison/plot.py

@@ -1,224 +1,180 @@
-###Plots##
-#Reordering the array so that compositional languages are at indices 0-3:
-c3a = [results3a[compIndices[0]],results3a[compIndices[1]], results3a[compIndices[2]],results3a[compIndices[3]]]
-c3b = [results3b[compIndices[0]],results3b[compIndices[1]], results3b[compIndices[2]],results3b[compIndices[3]]]
-c3c = [results3c[compIndices[0]],results3c[compIndices[1]], results3c[compIndices[2]],results3c[compIndices[3]]]
+from math import sqrt
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set_context(rc={'lines.markeredgewidth': 0.5})
+import csv
+import numpy as np
+import sys
 
-for i in range(len(compIndices)): 
-    results3a = np.delete(results3a,compIndices[i])
-    results3a = np.insert(results3a,i,c3a[i])
 
-    results3b = np.delete(results3b,compIndices[i])
-    results3b = np.insert(results3b,i,c3b[i])
+f1 = csv.reader(open('./results/comparison-il-samp-mean-a0-e0.050000-k3-g100-r10.csv'))
+f2 = csv.reader(open('./results/comparison-il-map-mean-a0-e0.050000-k3-g100-r10.csv'))
+f3 = csv.reader(open('./results/comparison-unwgh-samp-mean-a0-e0.050000-k3-g100-r10.csv'))
+f4 = csv.reader(open('./results/comparison-unwgh-map-mean-a0-e0.050000-k3-g100-r10.csv'))
+f5 = csv.reader(open('./results/comparison-wgh-samp-mean-a0-e0.050000-k3-g100-r10.csv'))
+f6 = csv.reader(open('./results/comparison-wgh-map-mean-a0-e0.050000-k3-g100-r10.csv'))
+f7 = csv.reader(open('./results/prior-a0-e0.050000-k3-g100-r10.csv'))
 
-    results3c = np.delete(results3c,compIndices[i])
-    results3c = np.insert(results3c,i,c3c[i])
+compIndices = [45, 57, 108, 120]
+
+def results_in_list(f):
+    results = []
+    firstline = True
+    for i in f:
+        if firstline:
+            firstline = False
+            continue
+        results.append(float(i[1]))
+    return results
+
+il_samp = results_in_list(f1)
[…]

### 2016-06-15 plot.py

@@ -1,68 +1,0 @@
-#####
-#Read in CSV, plot results.
-#####
-import numpy as np
-import matplotlib.pyplot as plt
-import seaborn as sns
-sns.set_context(rc={'lines.markeredgewidth': 0.5})
-import csv
-import sys
-
-f1 = csv.reader(open('./results/singlescalar-wgh-mean-a1-c0.900000-l30-k3-g30-r100.csv','rt'))
-f2 = csv.reader(open('./results/multiscalar-wgh-mean-a1-c0.900000-l30-k3-g30-r100.csv','rt'))
-
-def results_by_lexica_multi(f):
-    p = np.zeros(6)
-    firstline = True
-    for i in f:
-        if firstline:
-            firstline = False
-            continue
-        p[int(i[1])] += float(i[-1]) /3.
-        p[int(i[2])] += float(i[-1]) /3.
-        p[int(i[3])] += float(i[-1]) /3.
-    return p
-
-def results_by_lexica_single(f):
-    p = np.zeros(6)
-    firstline = True
-    for i in f:
-        if firstline:
-            firstline = False
-            continue
-        p[int(i[1])] += float(i[-1])
-    return p
-
-Y_single = results_by_lexica_single(f1)
-Y_multi = results_by_lexica_multi(f2)
-
-print Y_single
-print Y_multi
-###Plots##
-
-X = np.arange(6) #number of lexica
-
-
-
-
-width = 0.35
-#ymin,ymax = 0,np.max(np.array([Ya[np.argmax(results3a)],Yb[np.argmax(results3b)],Yc[np.argmax(results3c)]]))+0.01
-#ax.set_xticks(np.array([len(compIndices)/2,(len(hypotheses)-len(compIndices))/2]))
-
-fig, ax = plt.subplots()
-bar1 = ax.bar(X, Y_single, width,color='r')
-bar2 = ax.bar(X+width,Y_multi,width,color='y')
-
-plt.ylabel('Mean in population')
-ax.set_xticks(X+width)
-ax.set_xticklabels(('L1', 'L2', 'L3', 'L4','L5','L6'))
-ax.legend((bar1[0], bar2[0]), ('single scalar', '3 scalars'))
-ax.margins(0.025, 0.025)
-#ax.text(.5,.9,'$\alpha$ = 1, c = %r, $\lambda$ = %r, pairs = %r, k = %r',
-#        horizontalalignment='center',
[…]

### 2016-06-15 rmd-test-optimization.py

@@ -1,297 +1,0 @@
-#####
-#RMD for unweighted and weighted (community) learning with Bayesian (iterated) learning as M. 
-#3 pairs of scalar items, six lexica per scalar pair. 
-#2 possible signaling behaviors: literal or gricean
-#12 types (6 lexica * 2 signaling behaviors) per signaling pair
-#6*6*6 = 216  literal types, 216 gricean types, 432 types in total
-#####
-
-
-import numpy as np
-#np.set_printoptions(threshold=np.nan)
-from random import sample
-from itertools import product, permutations, combinations_with_replacement
-from player import LiteralPlayer,GriceanPlayer
-import sys 
-import datetime
-import csv
-
-alpha = 1 # rate to control difference between semantic and pragmatic violations
-cost = 0.9 # cost for LOT-concept with upper bound
-lam = 30 # soft-max parameter
-k = 3  # number of learning observations
-mc = 1000 #amount of samples from OBS when k > 4
-
-gens = 30 #number of generations per simulation run
-runs = 100 #number of independent simulation runs
-
-
-lexical_pairs = 3 #number of 2x2 matrices
-states = 2 #number of states
-messages = 2 #number of messages
-
-
-
-print '#Starting, ', datetime.datetime.now()
-lexica = [np.array( [[0.,0.],[1.,1.]] ), np.array( [[1.,1.],[0.,0.]] ), np.array( [[1.,1.],[1.,1.]] ), np.array( [[0.,1.],[1.,0.]] ), np.array( [[0.,1.],[1.,1.]] ), np.array( [[1.,1.],[1.,0.]] )]
-
-def types(l):
-    """outputs list of types from list of lexica input.""" 
-    typeList = []
-    for i in l:
-        typeList.append(LiteralPlayer(alpha,i))
-    for i in l:
-        typeList.append(GriceanPlayer(alpha,lam,i))
-    return typeList 
-
-typeList = types(lexica)
-
-
-def lex_prior(l):
-    """outputs learning prior vector from list of lexica input. Assumes twice as many types than lexica. First half of the vector is literal, second half is gricean."""
-    cost_v = np.zeros(len(lexica))
-    for j in range(len(l)):
-        lexicon_score = 0.
-        for i in range(l[j].shape[1]):
-            if (i == 0 or i == 1) and l[j][0,i] > 0 and l[j][1,i] == 0:
[…]

### 2016-06-21 paper/00specification.tex
clean up, observations by sampling and parametrized learning for single scalar pair

@@
 \section{Short exposition}
-The plan is to expand our work on the semantics/pragmatics divide by means of a (more extended) study on the lack of upper-bounds in the literal meaning of scalar expressions.
-The main idea of this extension, other than providing a more detailed treatment as well as a comparison with iterated learning, is to look at the effect multiple scalar pairs have on the dynamics.
-That is, we consider larger language fragments consisting of more than one pair of expressions of the ``some but not all'' vs.
-``all''-form.
+The plan is to expand our work on the semantics/pragmatics divide by means of a more involved \& principled study on the lack of upper-bounds in the literal meaning of scalar expressions.
+The main idea of this extension, other than providing a more detailed treatment as well as a comparison with iterated learning, is to do away with the assumption of communal learning.
+Potential extensions include more scalar pairs, the addition of an inductive bias towards uniform lexica across scalar pairs, and a variation of state frequencies.
 
 \section{Differences to CogSci setup.}
-\subparagraph{(I) Lexica and types.} $3$ pairs of scalar expressions.
-Each pair is a $(2,2)$-Boolean matrix, which means that a lexicon is now a $(8,8)$-matrix.
-We have two types of signaling behavior, literal and Gricean, and $6$ different $(2,2)$-lexica.
-The latter is the set of possible semantics of a single scalar pair.
-With three scalar pairs we have $6 \cdot 6 \cdot 6 = 216$ $(8,8)$-lexica.
-$216$ times two signaling behaviors makes for a total of $432$ types.
+\subparagraph{(I) No communal learning.} Only parental learning is considered
 
-\subparagraph{(II) Observations.} Before, the set of all production observations was $O =$\linebreak  $\{\tuple{\tuple{s_1,m_i},\tuple{s_2,m_j}} | m_i, m_j \in M\}$.
-A member of $O$ encodes that a teacher produced $m_i$ in state $s_1$ and $m_j$ in $s_2$, i.e.
-it encodes one witnessed message for each state.
+\subparagraph{(II) Sequences and atomic observations.} Before, the set of all observations was $O =$\linebreak  $\{\tuple{\tuple{s_1,m_i},\tuple{s_2,m_j}} | m_i, m_j \in M\}$.
+A member of $O$ encodes that a teacher produced $m_i$ in state $s_1$ and $m_j$ in $s_2$, i.e., it encodes one witnessed message for each state.
 A datum $d$ was a sequence of length $k$ of members of $O$.
@@
 
-\subparagraph{(III) Computational load.} Computing the $Q$-matrix from the set of all observation sequences of length $k$ is computationally intractable for large $k$.
-Here, we follow \citet{griffiths+kalish:2007} as well in sampling $1000$ sequences $d$ from the set of all sequences when $k > 4$.
-We use this sampled subset to compute $Q$ instead of the full set of sequences\footnote{Thomas: Assuming this solves the problem.
-My computer can't handle the computation even with sampling but a better computer may.
-I'm still waiting on people from my project to address my request for computation time on our server.}
+\subparagraph{(III) Observations as production.} Instead of taking the space of all possible sequences of length $k$ into consideration, we take sample from $O$ $k$-times according to the production probabilities of each type; $P(o = \tuple{s,m} | t_i) = P(s) P(m|s,t_i)$.
+$n$ such $k$-length sequences are sampled for each type.
+As a consequence, the data used for computing $Q_i$ is not the same as that used for $j$ $(i \neq j)$.
 
-\subparagraph{(IV) Mutation/Learning matrix.} If the results warrant it, classical (unweighted) learning instead of communal (weighted) learning.\footnote{Thomas: We can save the results for both cases in one go and decide on this later.}
+\subparagraph{(IV) Parametrized learning} $Q_{ij} \propto [\sum_d P(d|t_i) P(t_j|d)]^l$, where $l =1$ corresponds to probability matching and, as $l$ increases towards infinity, to MAP.
 
-\paragraph{Further modifications we may want to keep in mind (but currently don't make use of).}
+\section{Further modifications we may want to keep in mind (but currently don't make use of)}
 \subparagraph{(I) Cost for pragmatic reasoning.} At least in the CogSci setup the effect of adding cost to pragmatic reasoning is unsurprising: High cost for pragmatic signaling lowers the prevalence of pragmatic types.
@@
 
+\subparagraph{(V) More scalar pairs.} Taking into consideration more than one scalar pair.
+Preliminary results suggest that this does not influence the results in any meaningful way without further additions, e.g.
+by (III).
+
+\subparagraph{(VI) More lexica.} Not necessary. Preliminary results suggest that considering more lexica has no noteworthy effect on the dynamics (tested with all possible 2x2 lexica).
+
+\subparagraph{(VII) State frequencies.} Variations on state frequencies. This may have an interesting interaction with (III).
+
+\subparagraph{(VIII) Reintroduction of communal learning.} One possibility: The probably $N_{ij}$ with which a child of $t_i$ adopts $t_j$ could be the weighted sum of $Q_{ij}$ (as before) and a vector we get from learning from all of the population: $L_j = \sum_d P(d | \vec{p})  P(t_j | d)$, where $P(d | \vec{p}) = \sum_{i} P(d | t_i)  \vec{p}_i$ is the probability of observing $d$ when learning from a random member of the present population distribution.
+
 \section{General setup}
@@
 
-We consider a population of players with two signaling strategies, literal and Gricean (level $0$ and $1$ below), each equipped with one of $216$ lexicons.
-This yields a total of $432$ distinct player types $t \in T$.
-However, each lexicon is composed of three independent lexical pairs.
-That is, while $|M| = |S| = 8$, we can decompose each lexicon into three independent $(2,2)$-matrices.
-A lexicon is a combination of three such matrices.
+We consider a population of players with two signaling behaviors, literal and Gricean (level $0$ and $1$ below), each equipped with one of $6$ lexicons.
+This yields a total of $12$ distinct player types $t \in T$.
+$|M| = |S| = 2$, i.e., a lexicon is a $(2,2)$-matrix.
 These are listed in Table \ref{tab:lexica}.
@@
 \end{tabular}
-\caption{{\footnotesize Set of possible $(2,2)$-matrices, i.e. possible semantics for a pair of scalar expressions.}}
+\caption{{\footnotesize Set of possible $(2,2)$-matrices, i.e., a lexicon.}}
 \label{tab:lexica}
@@
     \item $\Phi = \sum_i x_i f_i$
-    \item $Q_{ij} \propto \sum_d P(d|t_i) \; P(t_j|d)$, where $P(t_j|d) \propto p^\star(t_j) P(d|t_j)$ and $d$ is a sequence of observations of length $k$ of the form \tuple{\tuple{s_i,m_j}, ...
-\tuple{s_k, m_l}}.
-\end{itemize}
-
-\paragraph{Weighted (communal) learning vs. unweighted (parental) learning}
-
-\begin{itemize}
+    \item $Q_{ij} \propto [\sum_d P(d|t_i) \; P(t_j|d)]^l$, where $P(t_j|d) \propto p^\star(t_j) P(d|t_j)$, $d$ is a sequence of observations of length $k$ of the form \tuple{\tuple{s_i,m_j}, ...
+\tuple{s_k, m_l}}, and $l \geq 1$ is a learning parameter.
 	\item For parental learning (standard RMD): $\dot x_i = \sum_j Q_{ji} \frac{x_j f_j}{\Phi}$
-	\item For communal learning (our variant in CogSci): $\dot x_i = \sum_j N_{ji} \frac{x_j f_j}{\Phi}$, where $N_{ji} = x'_{i} Q_{ji}$ and $x'_i$ is the proportion of type $i$ in the previous generation.
 \end{itemize}
@@
   \item True state distribution ($P$). We currently assume that $P = \frac{1}{|S|}$ but it may be interesting to vary this
+  \item Learning parameter $l \geq 1$ with $1$ corresponding to probability matching, and MAP as $l$ approaches infinity
+  \item $n$ is the sample of sequences of observations of length $k$ sampled from the production probabilities of each type
+  \item Number of generations $g$
 \end{enumerate}
@@
 \paragraph{Procedural description.} The game is initialized with some arbitrary distribution over player types.
-At the game's onset we compute $Q$ once based on the set of sequences $D$.
-If $k > 4$, $1000$ sequences are sampled from $D$ and used to compute $Q$ (instead of the full set).
+At the game's onset we compute $Q$ once based on the sets  of sequences $D$ (one for each parent type).
 Replicator dynamics are computed based on the fitness of each type in the current population as usual.
-In the case of community learning, $Q$ is weighted by the previous population vector in each new generation.
-
-\paragraph{Presentation.} We have a large number of types, some of which are highly similar.
-For instance, a type that uses $L_4$, $L_5$ and $L_4$ for her respective three scalar pairs is a different type than one that uses $L_5$, $L_4$ and $L_4$.
-As our goal is to elucidate the role and prevalence of upper-bounds, we may summarize the results not by type but by lexica counts.
-For instance, if the previous two types where all we had this would mean that $L_4$'s proportion is roughly $.66$ and $L_5$'s $.33$ (irrespective of signaling behavior).
-An example is given below for communal learning with a single scalar pair against three scalar pairs ($\alpha = 1, c = .7, \lambda = 30, k = 3$):
-
-\begin{figure}[h] %htb for exact position
-	\centering
-	\resizebox{13cm}{!}{\includegraphics{../plots/wgh-single-vs-multi-k3-c70}}
-\end{figure}
+$Q$ is computed anew for each independent run (of $g$ generations) given that it depends on $D$, which is sampled from production probabilities.
 

### 2016-06-21 1rmd-singlescalar.py

@@ -1,0 +1,172 @@
+#####
+#RMD with parametrized iterated parental learning 
+#1 pairs of scalar items, six lexica per scalar pair. 
+#2 possible signaling behaviors: literal or gricean
+#12 types (6 lexica * 2 signaling behaviors) per signaling pair
+#6 literal types, 6 gricean types, 12 types in total
+#####
+
+import numpy as np
+np.set_printoptions(threshold=np.nan)
+from random import sample
+from itertools import product
+from player import LiteralPlayer,GriceanPlayer
+import sys 
+import datetime
+import csv
+
+#####
+l1,l2,l3,l4,l5,l6 = np.array( [[0.,0.],[1.,1.]] ), np.array( [[1.,1.],[0.,0.]] ), np.array( [[1.,1.],[1.,1.]] ), np.array( [[0.,1.],[1.,0.]] ), np.array( [[0.,1.],[1.,1.]] ), np.array( [[1.,1.],[1.,0.]] )
+
+alpha = 1# rate to control difference between semantic and pragmatic violations
+cost = 0.1 # cost for LOT-concept with upper bound
+lam = 10 # soft-max parameter
+k = 3  # length of observation sequences
+sample_amount = 10 #amount of k-length samples for each production type
+
+gens = 20 #number of generations per simulation run
+runs = 50 #number of independent simulation runs
+
+states = 2 #number of states
+messages = 2 #number of messages
+
+learning_parameter = 1 #prob-matching = 1, increments approach MAP
+state_freq = np.ones(states) / float(states) #frequency of states s_1,...,s_n 
+
+
+f_mean = csv.writer(open('./results/singlescalar-mean-a%d-c%.2f-l%d-k%d-samples%d-learn%.2f-g%d-r%d.csv' %(alpha,cost,lam,k,sample_amount, learning_parameter, gens,runs),'wb')) #file to store mean results
+######
+
+
+
+print '#Starting, ', datetime.datetime.now()
+
+t1,t2,t3,t4,t5,t6 = LiteralPlayer(alpha,l1), LiteralPlayer(alpha,l2), LiteralPlayer(alpha,l3), LiteralPlayer(alpha,l4), LiteralPlayer(alpha,l5), LiteralPlayer(alpha,l6)
[…]

### 2016-07-05 paper/paper.tex
fragments for paper

--- 2016-06-21 paper/00specification.tex
@@
-\documentclass[11pt]{article}
+\documentclass[a4paper]{article}
 
-\usepackage[backgroundcolor=white,textsize=small,linecolor=red]{todonotes}
-
-\usepackage[margin=30mm]{geometry}
+\usepackage{geometry}
 \usepackage{natbib}
@@
 %--------------------
-% Math
-\usepackage[fleqn]{amsmath}
+\usepackage{gb4e}
+\noautomath
+
+\usepackage{amsmath}
 \usepackage{amsfonts}
@@
 \usepackage{amssymb}
-\usepackage{stmaryrd}
-\usepackage{multicol}
-\usepackage{graphicx}
-
-\newcommand{\mvalueof}[1]{\llbracket#1\rrbracket}
+%\usepackage{stmaryrd}
+%\usepackage{multicol}
+%\usepackage{graphicx}
+\usepackage{color}
+%\newcommand{\mvalueof}[1]{\llbracket#1\rrbracket}
 \newcommand{\citeposs}[2][]{\citeauthor{#2}'s (\citeyear[#1]{#2})}
@@
 
-\newcommand{\hloranj}[1]{\textcolor[rgb]{.8,.33,.0}{#1}}% prints in orange
-\newcommand{\argmax}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{max}}\;}
-\newcommand{\argmin}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{min}}\;}
-\newcommand{\sbna}{\exists\lnot\forall}
+\newcommand{\hl}[1]{\textcolor[rgb]{.8,.33,.0}{#1}}% prints in orange
+%\newcommand{\argmax}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{max}}\;}
+%\newcommand{\argmin}[1]{\underset{#1}{\operatorname{arg}\,\operatorname{min}}\;}
+%\newcommand{\sbna}{\exists\lnot\forall}
 
 %--------------------
-%\usepackage{gb4e}
 %
@@
 %
-%\usepackage{soul} %underlining
-%\setuldepth{a}
-
-%\usepackage{inconsolata}
-
 %-------------------
@@
 \author{%\bf NAME1 and NAME2\\
-    \today
+    ( -- draft \today --- )
 }
@@
 \maketitle
+\tableofcontents
+\section{Introduction}
+While there is disagreement on where its boundaries lie, a distinction between semantics and pragmatics is commonly agreed upon by linguists.
+Broadly put, semantics can be viewed as concerning the literal, truth-conditional, meaning of expressions whereas pragmatics concerns information conveyed by expressions that goes beyond its literal meaning.
+Consider, for instance, the following sentence
 
-\section{Short exposition}
-The plan is to expand our work on the semantics/pragmatics divide by means of a more involved \& principled study on the lack of upper-bounds in the literal meaning of scalar expressions.
-The main idea of this extension, other than providing a more detailed treatment as well as a comparison with iterated learning, is to do away with the assumption of communal learning.
-Potential extensions include more scalar pairs, the addition of an inductive bias towards uniform lexica across scalar pairs, and a variation of state frequencies.
+\begin{exe}
+\ex SI sentence \begin{xlist}
+	\ex Numeral n-1
+	\ex SI
+\end{xlist}
+\end{exe}
 
-\section{Differences to CogSci setup.}
-\subparagraph{(I) No communal learning.} Only parental learning is considered
+%Note first, that the literal meaning of {\em NUMERAL} is compatible with that of more informative or ``stronger'' alternatives such as {\em n+1, n+1} and so on.
 
+\paragraph{Semantics-pragmaits distinction and where the puzzle arises.}
+
+\paragraph{Description of plan to lay out model and test it on upper-bounds.}
+
+\section{Cultural transmission \& linguistic pressures}
+
+\paragraph{Overview of generally assumed linguistic pressures.} \hl{Focus has been put mainly on one over the other, even though it is mutually accepted that both are important.
+We involve both and connect this research to the rational probabilistic interlocutors literature}
+
+\subsection{Probabilistic models of language use}
+
+\paragraph{Language (fragments).}
+
+\paragraph{Signaling behavior.} With $\lambda \geq 1$ (rationality parameter), $\alpha \in [0,1]$ (pragmatic violations) and $pr \in \Delta(S)$ a common prior over $S$ (uniform so far):
+
+\begin{flalign}
+&R_{0}(s|m;L) \propto pr(s) L_{sm}\label{litl}\\
+&S_{0}(m|s;L) \propto \exp(\lambda \; L_{sm}) \label{lits}\\
+&R_{1}(s|m;L) \propto pr(s) S_{0}(m|s;L) \label{pragl}\\
+&S_{1}(m|s;L) \propto  \exp(\lambda \; R_{0}(s|m;L)^\alpha) \label{prags}
+\end{flalign}
+
+\subsection{Expressiveness}
+
+\paragraph{Symmetrized expected utility.} With $P \in \Delta(S)$ (uniform so far; therefore $P = pr$):
+\begin{itemize}
+
+  \item $U(t_i,t_j) = [U_S(t_i,t_j) + U_R(t_i,t_j)] / 2$
+  \item $U_S(t_i,t_j) = \sum_s P(s) \; \sum_m P_S(m|s;t_i) \; \sum_{s'} P_R(s'|m,t_j) \; \delta(s,s')$, where $\delta(s,s')$ returns $1$ iff $s = s'$ and otherwise $0$
+  \item $U_R(t_i,t_j) = U_S(t_j,t_i)$
+\end{itemize}
+
+\subsection{Learnability}
 \subparagraph{(II) Sequences and atomic observations.} Before, the set of all observations was $O =$\linebreak  $\{\tuple{\tuple{s_1,m_i},\tuple{s_2,m_j}} | m_i, m_j \in M\}$.
@@
 
-\subparagraph{(IV) Parametrized learning} $Q_{ij} \propto [\sum_d P(d|t_i) P(t_j|d)]^l$, where $l =1$ corresponds to probability matching and, as $l$ increases towards infinity, to MAP.
+\subparagraph{(IV) Parametrized learning} $Q_{ij} \propto \sum_d P(d|t_i) F(t_j,d)$, where $F(t_j,d) \propto P(t_j|d)^l$ and $l =1$ corresponds to probability matching and, as $l$ increases towards infinity, to MAP.
 
-\section{Further modifications we may want to keep in mind (but currently don't make use of)}
+The proportion of players of type $i$, $x_i$, is initialized as an arbitrary distribution over $T$.
+$p^\star \in \Delta(T)$ is learning a prior over (player) types dependent only on the lexicon of the type.
+\begin{itemize}
+    \item $f_i = \sum_j x_j U(x_i,x_j)$
+    \item $\Phi = \sum_i x_i f_i$
+    \item $Q_{ij} \propto [\sum_d P(d|t_i) \; P(t_j|d)]^l$, where $P(t_j|d) \propto p^\star(t_j) P(d|t_j)$, $d$ is a sequence of observations of length $k$ of the form \tuple{\tuple{s_i,m_j}, ...
+\tuple{s_k, m_l}}, and $l \geq 1$ is a learning parameter.
+	\item For parental learning (standard RMD): $\dot x_i = \sum_j Q_{ji} \frac{x_j f_j}{\Phi}$
+\end{itemize}
+
+\subsection{Summary}
+\paragraph{Procedural description.} The game is initialized with some arbitrary distribution over player types.
+At the game's onset we compute $Q$ once based on the sets  of sequences $D$ (one for each parent type).
+Replicator dynamics are computed based on the fitness of each type in the current population as usual.
+$Q$ is computed anew for each independent run (of $g$ generations) given that it depends on $D$, which is sampled from production probabilities.
+
+\section{Upper-bounds in weak scalar alternatives}
+\hl{phenomenon description}
+
+\subsection{Analysis}
+
+\paragraph{Languages.} We consider a population of players with two signaling behaviors, literal and Gricean (level $0$ and $1$ below), each equipped with one of $6$ lexicons.
+This yields a total of $12$ distinct player types $t \in T$.
+$|M| = |S| = 2$, i.e., a lexicon is a $(2,2)$-matrix.
+These are listed in Table \ref{tab:lexica}.
+
+\begin{table}[h]
+\centering 
+\begin{tabular}{l c l}
+$L_1$ = $\begin{pmatrix} 0 & 0 \\ 1 & 1 \end{pmatrix}$ & 
+$L_2$ = $\begin{pmatrix} 1 & 1 \\ 0 & 0 \end{pmatrix}$ & 
+$L_3$ = $\begin{pmatrix} 1 & 1 \\ 1 & 1 \end{pmatrix}$\\[0.5cm]
+
+$L_4$ = $\begin{pmatrix} 0 & 1 \\ 1 & 0 \end{pmatrix}$ &
+$L_5$ = $\begin{pmatrix} 0 & 1 \\ 1 & 1 \end{pmatrix}$ &
+$L_6$ = $\begin{pmatrix} 1 & 1 \\ 1 & 0 \end{pmatrix}$
+\end{tabular}
+\caption{{\footnotesize Set of possible $(2,2)$-matrices, i.e., a lexicon.}}
+\label{tab:lexica}
+\end{table}
+
+As in the CogSci paper, $L_4$ (semantic upper-bound for $m_2$) and $L_5$ (no semantic upper-bound for $m_2$) are the target lexica.
+Gricean $L_5$ users can convey/infer the bound pragmatically, while literal/Gricean $L_4$ users do so semantically.
+
+\subsection{Model parameters \& procedure} 
+\begin{enumerate}
+  \item Sequence length $k$
+  \item Pragmatic production parameter $\alpha$
+  \item Rationality parameter $\lambda$
+  \item Learning prior over types (lexica); cost parameter $c$.
+$p^\star(t_i) \propto n - c \cdot r$ where $n$ is the total number of states and $r$ that of upper-bounded messages only true of $s_1$ in $t_i$'s lexicon (if only $s_1$ is true of a message, then this message encodes an upper-bound).
+Then the score for $L_1$, $L_3$, $L_5$ is $2$, that of $L_4$ and $L_6$ is $2-c$, and that of $L_2$ is $2-2c$; Normalization over lexica scores yields the prior over lexica (which is equal to the prior over types).
+  \item Prior over meanings ($pr$). We assume that $pr(s) = \frac{1}{|S|}$ for all $s$.
+  \item True state distribution ($P$). We currently assume that $P = \frac{1}{|S|}$ but it may be interesting to vary this
+  \item Learning parameter $l \geq 1$ with $1$ corresponding to probability matching, and MAP as $l$ approaches infinity
+  \item $n$ is the sample of sequences of observations of length $k$ sampled from the production probabilities of each type
+  \item Number of generations $g$
+\end{enumerate}
+
+\section{Discussion}
 \subparagraph{(I) Cost for pragmatic reasoning.} At least in the CogSci setup the effect of adding cost to pragmatic reasoning is unsurprising: High cost for pragmatic signaling lowers the prevalence of pragmatic types.
@@
 
-\section{General setup}
-\subsection{Part 1: Signaling.}
-
-We consider a population of players with two signaling behaviors, literal and Gricean (level $0$ and $1$ below), each equipped with one of $6$ lexicons.
-This yields a total of $12$ distinct player types $t \in T$.
-$|M| = |S| = 2$, i.e., a lexicon is a $(2,2)$-matrix.
-These are listed in Table \ref{tab:lexica}.
-
-\begin{table}[h]
-\centering 
-\begin{tabular}{l c l}
-$L_1$ = $\begin{pmatrix} 0 & 0 \\ 1 & 1 \end{pmatrix}$ & 
-$L_2$ = $\begin{pmatrix} 1 & 1 \\ 0 & 0 \end{pmatrix}$ & 
-$L_3$ = $\begin{pmatrix} 1 & 1 \\ 1 & 1 \end{pmatrix}$\\[0.5cm]
-
-$L_4$ = $\begin{pmatrix} 0 & 1 \\ 1 & 0 \end{pmatrix}$ &
-$L_5$ = $\begin{pmatrix} 0 & 1 \\ 1 & 1 \end{pmatrix}$ &
-$L_6$ = $\begin{pmatrix} 1 & 1 \\ 1 & 0 \end{pmatrix}$
-\end{tabular}
-\caption{{\footnotesize Set of possible $(2,2)$-matrices, i.e., a lexicon.}}
-\label{tab:lexica}
-\end{table}
-
-As in the CogSci paper, $L_4$ (semantic upper-bound for $m_2$) and $L_5$ (no semantic upper-bound for $m_2$) are the target lexica.
-Gricean $L_5$ users can convey/infer the bound pragmatically, while literal/Gricean $L_4$ users do so semantically.
-
-\paragraph{Signaling behavior.} With $\lambda \geq 1$ (rationality parameter), $\alpha \in [0,1]$ (pragmatic violations) and $pr \in \Delta(S)$ a common prior over $S$ (uniform so far):
-
-\vspace{-0.15cm}
-\begin{flalign}
-&R_{0}(s|m;L) \propto pr(s) L_{sm}\label{litl}\\
-&S_{0}(m|s;L) \propto \exp(\lambda \; L_{sm}) \label{lits}\\
-&R_{1}(s|m;L) \propto pr(s) S_{0}(m|s;L) \label{pragl}\\
-&S_{1}(m|s;L) \propto  \exp(\lambda \; R_{0}(s|m;L)^\alpha) \label{prags}
-\end{flalign}
-
-\paragraph{Symmetrized expected utility.} With $P \in \Delta(S)$ (uniform so far; therefore $P = pr$):
-\begin{itemize}
-
-  \item $U(t_i,t_j) = [U_S(t_i,t_j) + U_R(t_i,t_j)] / 2$
-  \item $U_S(t_i,t_j) = \sum_s P(s) \; \sum_m P_S(m|s;t_i) \; \sum_{s'} P_R(s'|m,t_j) \; \delta(s,s')$, where $\delta(s,s')$ returns $1$ iff $s = s'$ and otherwise $0$
-  \item $U_R(t_i,t_j) = U_S(t_j,t_i)$
-\end{itemize}
-
-\subsection{Part 2: Mutation \& replication.}
-The proportion of players of type $i$, $x_i$, is initialized as an arbitrary distribution over $T$.
-$p^\star \in \Delta(T)$ is learning a prior over (player) types dependent only on the lexicon of the type.
-\begin{itemize}
-    \item $f_i = \sum_j x_j U(x_i,x_j)$
-    \item $\Phi = \sum_i x_i f_i$
-    \item $Q_{ij} \propto [\sum_d P(d|t_i) \; P(t_j|d)]^l$, where $P(t_j|d) \propto p^\star(t_j) P(d|t_j)$, $d$ is a sequence of observations of length $k$ of the form \tuple{\tuple{s_i,m_j}, ...
-\tuple{s_k, m_l}}, and $l \geq 1$ is a learning parameter.
-	\item For parental learning (standard RMD): $\dot x_i = \sum_j Q_{ji} \frac{x_j f_j}{\Phi}$
-\end{itemize}
-
-\subsection{Model parameters \& procedure} 
-\begin{enumerate}
-  \item Sequence length $k$
-  \item Pragmatic production parameter $\alpha$
-  \item Rationality parameter $\lambda$
-  \item Learning prior over types (lexica); cost parameter $c$.
-$p^\star(t_i) \propto n - c \cdot r$ where $n$ is the total number of states and $r$ that of upper-bounded messages only true of $s_1$ in $t_i$'s lexicon (if only $s_1$ is true of a message, then this message encodes an upper-bound).
-Then the score for $L_1$, $L_3$, $L_5$ is $2$, that of $L_4$ and $L_6$ is $2-c$, and that of $L_2$ is $2-2c$; Normalization over lexica scores yields the prior over lexica (which is equal to the prior over types).
-  \item Prior over meanings ($pr$). We assume that $pr(s) = \frac{1}{|S|}$ for all $s$.
-  \item True state distribution ($P$). We currently assume that $P = \frac{1}{|S|}$ but it may be interesting to vary this
-  \item Learning parameter $l \geq 1$ with $1$ corresponding to probability matching, and MAP as $l$ approaches infinity
-  \item $n$ is the sample of sequences of observations of length $k$ sampled from the production probabilities of each type
-  \item Number of generations $g$
-\end{enumerate}
-
-\paragraph{Procedural description.} The game is initialized with some arbitrary distribution over player types.
-At the game's onset we compute $Q$ once based on the sets  of sequences $D$ (one for each parent type).
-Replicator dynamics are computed based on the fitness of each type in the current population as usual.
-$Q$ is computed anew for each independent run (of $g$ generations) given that it depends on $D$, which is sampled from production probabilities.
+\section{Conclusion}
 

### 2016-07-06 1rmd-singlescalar.py
minor edit code in 1rmd-singlescalar.py

@@ -18,7 +18,7 @@ import csv
 #####
 l1,l2,l3,l4,l5,l6 = np.array( [[0.,0.],[1.,1.]] ), np.array( [[1.,1.],[0.,0.]] ), np.array( [[1.,1.],[1.,1.]] ), np.array( [[0.,1.],[1.,0.]] ), np.array( [[0.,1.],[1.,1.]] ), np.array( [[1.,1.],[1.,0.]] )
 
-alpha = 1# rate to control difference between semantic and pragmatic violations
+alpha = 1 # rate to control difference between semantic and pragmatic violations
 cost = 0.1 # cost for LOT-concept with upper bound
 lam = 10 # soft-max parameter
 k = 3  # length of observation sequences
@@ -101,9 +101,13 @@ def get_mutation_matrix(k):
         lhs = get_likelihood(type_obs) #P(parent data|t_i) for all types
         post = normalize(lexica_prior * np.transpose(lhs)) #P(t_j|parent data) for all types; P(d|t_j)P(t_j)
         parametrized_post = normalize(post**learning_parameter)
-        parent_production_q = np.dot(lhs,parametrized_post)
-        
-        out[parent_type] = parent_production_q[parent_type] #store the parents row Q_{parent_type}
+
+        ## TB original:
+        #parent_production_q = np.dot(lhs,parametrized_post)
+        #out[parent_type] = parent_production_q[parent_type] #store the parents row Q_{parent_type}
+
+        ## MF alternative:
+        out[parent_type] = np.dot(np.transpose(lhs[parent_type]),parametrized_post)
 
     return normalize(out)
 

### 2016-07-13 paper/paper.tex
data subset

@@
 \usepackage{natbib}
+\bibpunct[,]{(}{)}{,}{a}{}{;}
 
@@
 \tableofcontents
-\section{Introduction}
-While there is disagreement on where its boundaries lie, a distinction between semantics and pragmatics is commonly agreed upon by linguists.
-Broadly put, semantics can be viewed as concerning the literal, truth-conditional, meaning of expressions whereas pragmatics concerns information conveyed by expressions that goes beyond its literal meaning.
-Consider, for instance, the following sentence
+\section{Introduction}\label{sec:introduction}
+
+\hl{some general sentence here}
+In linguistic theorizing, it is common to draw a distinction between semantics and pragmatics.
+Broadly speaking, semantics concerns the truth-conditional content of expressions, whereas pragmatics deals with information beyond such literal meanings.
+For instance, the integration of context and the derivation of defeasible inferences.
+Thus, the information conveyed is seldom if ever solely determined by semantics, but rather by its pragmatic enrichment.
+The present investigation addresses the selection of particular semantics over others in light of pragmatics.
+%  the semantics-pragmatics distinction raises the challenge of justifying the former's structure.
+
+Consider, for instance, the utterance {\em Bill read five books}.
+Truth-conditionally, this utterance does not set an upper-bound for the amount of books that Bill read; he may have read six, seven, or more books.
+Pragmatically, however, an addressee who assumes the speaker to be cooperative and knowledgeable may draw the (defeasible) inference that {\em Bill read no more than five books}.
+After all, had he read more, for example, six books, and the speaker had known this, she would have said {\em Bill read six books} instead.
+In this way, what is conveyed (pragmatically) goes beyond what is (literally) said \citep{grice:1975}.
+
+Much research in semantics and pragmatics has been aimed at characterizing the information conveyed by classes of expressions, such as numerals, in terms of either domain, or their interplay.
+Thus, while there is disagreement on where their boundary lies, their distinction remains a core (if tacit) assumption in linguistics.
+An issue that has received little attention, however, is the justification of semantic structure in view of pragmatics.
+Prima facie, under a functionalist view, is is puzzling that {\em Bill read five books} does not semantically rule out its more informative or ``stronger'' alternatives.
+Wouldn't it serve speakers of a language better if weak(er) expressions such as {\em warm}, {\em or}, {\em some} or {\em big} were not truth-conditionally compatible with stronger alternatives such as {\em hot}, {\em and}, {\em all} and {\em huge}, respectively?
+After all, this is what an interlocutor's pragmatically enriched interpretation yields in the case of so-called scalar inferences.
+Once a distinction between semantics and pragmatics is drawn, similar questions can be posed for other types of pragmatic inferences, such as ignorance inferences as (\ref{ex:ignorance}), or manner inferences such as (\ref{ex:manner-1}) and (\ref{ex:manner-2}).
 
 \begin{exe}
-\ex SI sentence \begin{xlist}
-	\ex Numeral n-1
-	\ex SI
+\ex\label{ex:ignorance} \begin{xlist}
+	\ex Bill read at least five books
+	\ex $\leadsto$ Bill read five or more books (but the speaker doesn't know how many)
+\end{xlist}
+\ex \label{ex:manner-1} \begin{xlist}
+	\ex Mercader killed Trotsky.
+	\ex $\leadsto$ Mercader killed Trotsky in a stereotypical way
+    \end{xlist}
+\ex \label{ex:manner-2} \begin{xlist} \ex Mercader caused Trotsky to die
+		  \ex $\leadsto$ Mercader killed Trotsky in a non-stereotypical way
     \end{xlist}
@@
 
-%Note first, that the literal meaning of {\em NUMERAL} is compatible with that of more informative or ``stronger'' alternatives such as {\em n+1, n+1} and so on.
+In more general terms, the question then is what motivates the selection of particular semantic structures over other hypothetical alternatives in light of pragmatics.
+While a number of recent investigations have began to address issues pertaining to the development and selection of linguistic features, the semantics-pragmatics distinction has received little attention (see \citealt{steels:2015} and \citealt{tamariz+kirby:2016} for recent overviews).
+Our starting point is given by the overarching account of competing pressures that crystallizes across different frameworks: Natural languages need to allow for successful communication, one the one hand, and be well suited for acquisition on the other.
+Put differently, languages need to be well-adapted to the communicative needs within a linguistic community, and be learnable to survive their faithful transmission across generations.
 
-\paragraph{Semantics-pragmaits distinction and where the puzzle arises.}
+ \hl{Description of plan to lay out model and test it on upper-bounds.}
 
-\paragraph{Description of plan to lay out model and test it on upper-bounds.}
+\hl{general notes on cultural evolution, biological evolution, and their connection}
 
-\section{Cultural transmission \& linguistic pressures}
+\section{Linguistic structure \& selectional dynamics}
+The emergence and change of linguistic structure is driven by many factors, from biological and socio-ecological to cultural \citep{steels:2011,tamariz+kirby:2016}.
+Broadly put, social and ecological pressures determine communicative needs, while biology determines the architecture available for its use.
+Our focus is on the latter, cultural, factor, wherein linguistic structure is analyzed in terms of its use, as well as its transmission across generations.
 
-\paragraph{Overview of generally assumed linguistic pressures.} \hl{Focus has been put mainly on one over the other, even though it is mutually accepted that both are important.
-We involve both and connect this research to the rational probabilistic interlocutors literature}
+As already mentioned in \S\ref{sec:introduction}, research on selectional forces that apply in the cultural evolution of language has focused on two main pressures: expressivity and learnability.
+However, while it is generally acknowledged that both play a pivotal role, past approaches have focused exclusively, or at least emphasized, the role of one over the other (a recent exception is \citealt{kirby+etal:2015}).
+
+Expressiveness, or communicative efficiency, has been at the center of applications of evolutionary game theory to linguistics \citep{nowak+krakauer:1999,huttegger+zollman:2013},  \hl{explain RMD}
+
+as well as \hl{Luc Steels stuff}.
+
+In contrast, the iterated learning paradigm has focused on the effects of language transmission from generations of speakers to the next. \hl{explain IL}
 

### 2016-07-13 1rmd-singlescalar.py

@@ -34,7 +34,10 @@ learning_parameter = 1 #prob-matching = 1, increments approach MAP
 state_freq = np.ones(states) / float(states) #frequency of states s_1,...,s_n 
 
 
-f_mean = csv.writer(open('./results/singlescalar-mean-a%d-c%.2f-l%d-k%d-samples%d-learn%.2f-g%d-r%d.csv' %(alpha,cost,lam,k,sample_amount, learning_parameter, gens,runs),'wb')) #file to store mean results
+f = csv.writer(open('./results/singlescalar-a%d-c%.2f-l%d-k%d-samples%d-learn%.2f-g%d-r%d.csv' %(alpha,cost,lam,k,sample_amount, learning_parameter, gens,runs),'wb')) #file to store mean results
+
+f.writerow(["run_ID", "t1_initial", "t2_initial","t3_initial","t4_initial","t5_initial","t6_initial","t7_initial","t8_initial","t9_initial","t10_initial","t11_initial","t12_initial","alpha", "prior_cost_c", "lambda", "k", "sample_amount", "learning_parameter", "generations", "t1_final", "t2_final","t3_final","t4_final","t5_final","t6_final","t7_final","t8_final","t9_final","t10_final","t11_final","t12_final"])
+
 ######
 
 
@@ -102,11 +105,6 @@ def get_mutation_matrix(k):
         post = normalize(lexica_prior * np.transpose(lhs)) #P(t_j|parent data) for all types; P(d|t_j)P(t_j)
         parametrized_post = normalize(post**learning_parameter)
 
-        ## TB original:
-        #parent_production_q = np.dot(lhs,parametrized_post)
-        #out[parent_type] = parent_production_q[parent_type] #store the parents row Q_{parent_type}
-
-        ## MF alternative:
         out[parent_type] = np.dot(np.transpose(lhs[parent_type]),parametrized_post)
 
     return normalize(out)
@@ -127,34 +125,25 @@ u = get_utils()
 ###Multiple runs
 print '#Beginning multiple runs, ', datetime.datetime.now()
 
-f_mean.writerow(["sig_behavior","lexicon","proportion"])
 p_sum = np.zeros(len(typeList)) #vector to store results from a run
 
 for i in xrange(runs):
     q = get_mutation_matrix(k) #compute new Q for each new run, as it now depends on sampled observations
[…]

### 2016-07-13 csv-reader.py

@@ -1,0 +1,113 @@
+#####
+#Read in CSV, plot results.
+#####
+import numpy as np
+import matplotlib.pyplot as plt
+import seaborn as sns
+sns.set_context(rc={'lines.markeredgewidth': 0.5})
+import csv
+import sys
+
+d = open('./results/singlescalar-a1-c0.10-l10-k3-samples10-learn1.00-g20-r50.csv','rt')
+f = csv.reader(d)
+
+def initial_population(f):
+    out = []
+    firstline = True
+    d.seek(0)
+    for i in f:
+        p = np.zeros(12)
+        if firstline:
+            firstline = False
+            continue
+        p[0] = float(i[1])
+        p[1] = float(i[2])
+        p[2] = float(i[3])
+        p[3] = float(i[4])
+        p[4] = float(i[5])
+        p[5] = float(i[6])
+        p[6] = float(i[7])
+        p[7] = float(i[8])
+        p[8] = float(i[9])
+        p[9] = float(i[10])
+        p[10] = float(i[11])
+        p[11] = float(i[12])
+        out.append(p)
+    return out
+
+def final_population(f):
+    out = []
+    firstline = True
+    d.seek(0)
+    for i in f:
+        p = np.zeros(12)
+        if firstline:
+            firstline = False
+            continue
+        p[0] = float(i[-12])
+        p[1] = float(i[-11])
+        p[2] = float(i[-10])
+        p[3] = float(i[-9])
+        p[4] = float(i[-8])
+        p[5] = float(i[-7])
+        p[6] = float(i[-6])
+        p[7] = float(i[-5])
+        p[8] = float(i[-4])
+        p[9] = float(i[-3])
+        p[10] = float(i[-2])
+        p[11] = float(i[-1])
+        out.append(p)
+    return out
+
+
+p_initial_list = initial_population(f)
+p_final_list = final_population(f)
+
+initial_mean = sum(p_initial_list) / len(p_initial_list)
+final_mean = sum(p_final_list) / len(p_final_list)
+
+
+sys.exit()
+def results_by_lexica_single(f):
+    p = np.zeros(6)
+    firstline = True
+    for i in f:
+        if firstline:
+            firstline = False
+            continue
[…]

### 2016-07-13 pd-csv.py

@@ -1,0 +1,185 @@
+###
+import pandas as pd
+import numpy as np 
+import matplotlib.pyplot as plt
+import seaborn as sns
+#sns.set_context(rc={'lines.markeredgewidth': 0.5})
+import os
+import glob
+###
+
+###Column headers:###
+#run_ID,t1_initial,t2_initial,t3_initial,t4_initial,t5_initial,t6_initial,t7_initial,t8_initial,t9_initial,t10_initial,t11_initial,t12_initial,
+#alpha,prior_cost_c,lambda,k,sample_amount,learning_parameter,generations,
+#t1_final,t2_final,t3_final,t4_final,t5_final,t6_final,t7_final,t8_final,t9_final,t10_final,t11_final,t12_final
+#####################
+
+
+print 'Loading data'
+path =r'./resultssubset' 
+all_files = glob.glob(os.path.join(path, "*.csv"))     # advisable to use os.path.join as this makes concatenation OS independent
+df = pd.concat(pd.read_csv(f) for f in all_files)
+
+
+###Plot 1###
+a = 1
+c = 0.2
+k = 5
+lam = 10
+sample = 15
+learn = 1
+
+final_group = df.loc[:,'alpha':]
+final_group = final_group.loc[final_group['alpha'] == a]
+final_group = final_group.loc[final_group['prior_cost_c'] == c]
+final_group = final_group.loc[final_group['k'] == k]
+final_group = final_group.loc[final_group['lambda'] == lam]
+final_group = final_group.loc[final_group['sample_amount'] == sample]
+#final_group = final_group.loc[final_group['learning_parameter'] == learn]
+
+
+all_finals = final_group.groupby(['learning_parameter'])
+sub_finals = all_finals[['t10_final','t11_final','t12_final']].agg(np.average)
+
+print sub_finals
+
+ax = sub_finals.plot(title='alpha = %d, $\lambda = %d$, c = %.2f,k = %d, samples = %d' %(a,lam,c,k,sample))
+ax.set(ylabel="proportion",xlabel='learning parameter')
+plt.legend(["$pragmatic_{10}$","$pragmatic_{11}$","$pragmatic_{12}$"], loc='best')
+
+plt.show()
+
+
+
+###Plot 2###
+a = 1
+c = 0.5
+k = 5
+lam = 10
+sample = 15
+learn = 1
+
+final_group = df.loc[:,'alpha':]
+final_group = final_group.loc[final_group['alpha'] == a]
[…]

### 2016-07-13 run-analysis.py

@@ -1,0 +1,19 @@
+import modsinglescalar as rmd
+#parameter order for run(*) is run(alpha,cost,lam,k,sample_amount, learning_parameter,gens,runs)
+
+alpha = [0.25,0.5,0.7,1]
+cost = [0.01,0.05,0.1,0.2,0.3,0.4,0.5,0.7,0.8,0.9]
+lam = [1,10,30,50]
+seq_length = [1,3,5,9,13,15,17,20]
+sample_amount = [1,3,5,7,9,12,15] + [20,40,60,80,100] #+ [200,300,400,500]
+learning_parameter = [1,2,4,6,8,10]
+gens = 20
+runs = 1000
+
+for a in alpha:
+    for c in cost:
+        for l in lam:
+            for k in seq_length:
+                for sampl in sample_amount:
+                    for learn in learning_parameter:
+                        rmd.run(a,c,l,k,sampl,learn,gens,runs)

### 2016-07-13 run-small.py

@@ -1,0 +1,18 @@
+import modsinglescalar as rmd
+#parameter order for run(*) is run(alpha,cost,lam,k,sample_amount, learning_parameter,gens,runs)
+
+a = 1
+lam = [10,30]
+cost = [0.01,0.05,0.1,0.2,0.3,0.4,0.5,0.7,0.8,0.9]
+seq_length = [1,3,5,9,13,15,17,20]
+sample_amount = [1,3,5,7,9,12,15]  #+ [200,300,400,500]
+learning_parameter = [1,2,4,6,8,10]
+gens = 20
+runs = 1000
+
+for c in cost:
+    for l in lam:
+        for k in seq_length:
+            for sampl in sample_amount:
+                for learn in learning_parameter:
+                    rmd.run(a,c,l,k,sampl,learn,gens,runs)

### 2016-07-13 1rmd-singlescalar.py
small changes in pandas-code

@@ -20,7 +20,7 @@ l1,l2,l3,l4,l5,l6 = np.array( [[0.,0.],[1.,1.]] ), np.array( [[1.,1.],[0.,0.]] )
 
 alpha = 1 # rate to control difference between semantic and pragmatic violations
 cost = 0.1 # cost for LOT-concept with upper bound
-lam = 10 # soft-max parameter
+lam = 30 # soft-max parameter
 k = 3  # length of observation sequences
 sample_amount = 10 #amount of k-length samples for each production type
 
@@ -54,6 +54,7 @@ likelihoods = np.array([t.sender_matrix for t in typeList])
 
 lexica_prior = np.array([2, 2- 2* cost, 2, 2 - cost , 2 , 2-cost, 2, 2- 2* cost, 2, 2 - cost , 2 , 2-cost])
 lexica_prior = lexica_prior / sum(lexica_prior)
+
 
 def normalize(m):
     return m / m.sum(axis=1)[:, np.newaxis]

### 2016-07-13 csv-reader.py

@@ -1,113 +1,0 @@
-#####
-#Read in CSV, plot results.
-#####
-import numpy as np
-import matplotlib.pyplot as plt
-import seaborn as sns
-sns.set_context(rc={'lines.markeredgewidth': 0.5})
-import csv
-import sys
-
-d = open('./results/singlescalar-a1-c0.10-l10-k3-samples10-learn1.00-g20-r50.csv','rt')
-f = csv.reader(d)
-
-def initial_population(f):
-    out = []
-    firstline = True
-    d.seek(0)
-    for i in f:
-        p = np.zeros(12)
-        if firstline:
-            firstline = False
-            continue
-        p[0] = float(i[1])
-        p[1] = float(i[2])
-        p[2] = float(i[3])
-        p[3] = float(i[4])
-        p[4] = float(i[5])
-        p[5] = float(i[6])
-        p[6] = float(i[7])
-        p[7] = float(i[8])
-        p[8] = float(i[9])
-        p[9] = float(i[10])
-        p[10] = float(i[11])
-        p[11] = float(i[12])
-        out.append(p)
-    return out
-
-def final_population(f):
-    out = []
-    firstline = True
-    d.seek(0)
-    for i in f:
-        p = np.zeros(12)
-        if firstline:
-            firstline = False
-            continue
-        p[0] = float(i[-12])
-        p[1] = float(i[-11])
-        p[2] = float(i[-10])
-        p[3] = float(i[-9])
-        p[4] = float(i[-8])
-        p[5] = float(i[-7])
-        p[6] = float(i[-6])
-        p[7] = float(i[-5])
-        p[8] = float(i[-4])
-        p[9] = float(i[-3])
-        p[10] = float(i[-2])
-        p[11] = float(i[-1])
-        out.append(p)
-    return out
-
-
-p_initial_list = initial_population(f)
-p_final_list = final_population(f)
-
-initial_mean = sum(p_initial_list) / len(p_initial_list)
-final_mean = sum(p_final_list) / len(p_final_list)
-
-
-sys.exit()
-def results_by_lexica_single(f):
-    p = np.zeros(6)
-    firstline = True
-    for i in f:
-        if firstline:
-            firstline = False
-            continue
[…]

### 2016-07-13 pd-csv.py

@@ -21,6 +21,7 @@ all_files = glob.glob(os.path.join(path, "*.csv"))     # advisable to use os.pat
 df = pd.concat(pd.read_csv(f) for f in all_files)
 
 
+sys.exit()
 ###Plot 1###
 a = 1
 c = 0.2

### 2016-07-14 pd-csv.py
data and subset analysis

@@ -6,6 +6,7 @@ import seaborn as sns
 #sns.set_context(rc={'lines.markeredgewidth': 0.5})
 import os
 import glob
+import sys
 ###
 
 ###Column headers:###
@@ -183,4 +184,37 @@ ax.set(ylabel="proportion",xlabel='sequence length k')
 plt.legend(["$pragmatic_{10}$","$pragmatic_{11}$","$pragmatic_{12}$"], loc='best')
 
 plt.show()
+
+###
+a = 1
+#c = 0.3
+k = 5
+lam = 30
+sample = 15
+#learn = 4
+
+final_group = df.loc[:,('alpha','prior_cost_c','lambda','k','sample_amount','learning_parameter','t11_final')]
+
+group = final_group.loc[final_group['alpha'] == a]
+#group = group.loc[group['prior_cost_c'] == c]
+group = group.loc[group['k'] == k]
+group = group.loc[group['lambda'] == lam]
+group = group.loc[group['sample_amount'] == sample]
+#group = group.loc[group['learning_parameter'] == learn]
+
+def cust_mean(grp):
+    grp['mean'] = grp['t11_final'].mean()
+    return grp
+
+t11_final = group.groupby(['prior_cost_c', 'learning_parameter']).apply(cust_mean) #add new column with mean
+t11_rel = t11_final.loc[:,('prior_cost_c', 'learning_parameter','mean')] #ignore other columns, given that they are fixed
+t11_rel_uniq = t11_rel.drop_duplicates() #drop duplictes that arise from adding the mean column
+t11_rec = t11_rel_uniq.pivot('prior_cost_c','learning_parameter','mean') #reshape to have a prior_cost_c by learning_parameter table
+
+sns.set(font_scale=1.2)
+ax = sns.heatmap(t11_rec) 
+ax.set(ylabel='prior parameter c',xlabel='posterior parameter l', title=r'Pragmatic $L_5$ ($\alpha = %d, \lambda = %d, samples = %d, k = %d$)' %(a,lam,sample,k))
+ax.invert_yaxis()
+plt.show()
+
 

### 2016-07-14 presentations/edindata/plots.py

@@ -1,0 +1,104 @@
+###
+import pandas as pd
+import numpy as np 
+import matplotlib.pyplot as plt
+import seaborn as sns
+#sns.set_context(rc={'lines.markeredgewidth': 0.5})
+import os
+import glob
+import sys
+###
+###Column headers:###
+#run_ID,t1_initial,t2_initial,t3_initial,t4_initial,t5_initial,t6_initial,t7_initial,t8_initial,t9_initial,t10_initial,t11_initial,t12_initial,
+#alpha,prior_cost_c,lambda,k,sample_amount,learning_parameter,generations,
+#t1_final,t2_final,t3_final,t4_final,t5_final,t6_final,t7_final,t8_final,t9_final,t10_final,t11_final,t12_final
+#####################
+
+
+print 'Loading data'
+path =r'./results1' 
+all_files = glob.glob(os.path.join(path, "*.csv"))     # advisable to use os.path.join as this makes concatenation OS independent
+df = pd.concat(pd.read_csv(f) for f in all_files)
+
+def cust_mean(grp):
+    grp['mean'] = grp['t11_final'].mean()
+    return grp
+
+
+### Heatplot prior/posterior ####
+a = 1
+k = 5
+lam = 30
+sample = 10
+
+final_group = df.loc[:,('alpha','prior_cost_c','lambda','k','sample_amount','learning_parameter','t11_final')]
+
+group = final_group.loc[final_group['alpha'] == a]
+group = group.loc[group['k'] == k]
+group = group.loc[group['lambda'] == lam]
+group = group.loc[group['sample_amount'] == sample]
+
+
+t11_final = group.groupby(['prior_cost_c', 'learning_parameter']).apply(cust_mean) #add new column with mean
+t11_rel = t11_final.loc[:,('prior_cost_c', 'learning_parameter','mean')] #ignore other columns, given that they are fixed
+t11_rel_uniq = t11_rel.drop_duplicates() #drop duplictes that arise from adding the mean column
+t11_rec = t11_rel_uniq.pivot('prior_cost_c','learning_parameter','mean') #reshape to have a prior_cost_c by learning_parameter table
+
+
+#yticks = np.arange(min(t11_rec.index),max(t11_rec.index),0.15)
+sns.set(font_scale=1.2)
[…]

### 2016-07-14 presentations/edindata/run-small.py

@@ -1,0 +1,32 @@
+import modsinglescalar as rmd
+#parameter order for run(*) is run(alpha,cost,lam,k,sample_amount, learning_parameter,gens,runs)
+
+
+def cost_to_learning():
+    a = 1
+    lam = 30
+    cost = [x/100. for x in xrange(100)]
+    seq_length = 5
+    sample_amount = 10
+    learning_parameter = [x for x in xrange(1,11)]
+    gens = 20
+    runs = 1000
+    
+    for c in cost:
+        for learn in learning_parameter:
+            rmd.run(a,c,lam,seq_length,sample_amount,learn,gens,runs)
+
+def ration_to_seq():
+    a = 1
+    lam = [1,10,20,30,40,50]
+    c = .4
+    seq_length = [x for x in xrange(1,20)]
+    sample_amount = 20
+    learn = 3
+    gens = 20
+    runs = 1000
+    
+    for l in lam:
+        for k in seq_length:
+            rmd.run(a,c,l,k,sample_amount,learn,gens,runs)
+

### 2016-07-16 presentations/edindata/plots.py
fixed plots

@@ -44,15 +44,19 @@ t11_rel = t11_final.loc[:,('prior_cost_c', 'learning_parameter','mean')] #ignore
 t11_rel_uniq = t11_rel.drop_duplicates() #drop duplictes that arise from adding the mean column
 t11_rec = t11_rel_uniq.pivot('prior_cost_c','learning_parameter','mean') #reshape to have a prior_cost_c by learning_parameter table
 
+sns.set(font_scale=1.2)
 
 #yticks = np.arange(min(t11_rec.index),max(t11_rec.index),0.15)
-sns.set(font_scale=1.2)
-ax = sns.heatmap(t11_rec)#, yticklabels=yticks) 
+vacio = ["" for _ in xrange(9)]
+yticks = [0] + vacio + [0.1] + vacio + [0.2] + vacio + [0.3] + vacio + [0.4] + vacio + [0.5] + vacio + [0.6] + vacio + [0.7] + vacio + [0.8] + vacio + [0.9] + ["" for _ in xrange(8)] + [0.99]
+ax = sns.heatmap(t11_rec,yticklabels=yticks)#, yticklabels=yticks) 
 ax.set(ylabel='Prior parameter c',xlabel='Posterior parameter l', title=r'Pragmatic $L5$ ($\alpha = %d, \lambda = %d$, samples = %d, k = %d)' %(a,lam,sample,k))
-#plt.yticks(arange(5), (0,2,4,6,8), rotation=0)
+plt.yticks(rotation=0)
+
 ax.invert_yaxis()
 plt.show()
 
+sys.exit()
 
 ### development over cost ###
 

### 2016-07-16 presentations/edindata/plots2.py

@@ -1,0 +1,161 @@
+###
+import pandas as pd
+import numpy as np 
+import matplotlib.pyplot as plt
+import seaborn as sns
+#sns.set_context(rc={'lines.markeredgewidth': 0.5})
+import os
+import glob
+import sys
+###
+###Column headers:###
+#run_ID,t1_initial,t2_initial,t3_initial,t4_initial,t5_initial,t6_initial,t7_initial,t8_initial,t9_initial,t10_initial,t11_initial,t12_initial,
+#alpha,prior_cost_c,lambda,k,sample_amount,learning_parameter,generations,
+#t1_final,t2_final,t3_final,t4_final,t5_final,t6_final,t7_final,t8_final,t9_final,t10_final,t11_final,t12_final
+#####################
+
+
+print 'Loading data'
+path =r'./results2' 
+all_files = glob.glob(os.path.join(path, "*.csv"))     # advisable to use os.path.join as this makes concatenation OS independent
+df = pd.concat(pd.read_csv(f) for f in all_files)
+
+def cust_mean(grp):
+    grp['mean'] = grp['t11_final'].mean()
+    return grp
+
+### Heatplot seq-length/lam ####
+a = 1
+sample = 10
+c = .4
+sample_amount = 20
+learn = 3
+gens = 20
+runs = 1000
+
+
+final_group = df.loc[:,('alpha','prior_cost_c','lambda','k','sample_amount','learning_parameter','t11_final')]
+
+group = final_group.loc[final_group['alpha'] == a]
+group = group.loc[group['prior_cost_c'] == c]
+group = group.loc[group['sample_amount'] == sample]
+group = group.loc[group['learning_parameter'] == learn]
+
+t11_final = group.groupby(['lambda','k']).apply(cust_mean) #add new column with mean
+t11_rel = t11_final.loc[:,('lambda','k','mean')] #ignore other columns, given that they are fixed
+t11_rel_uniq = t11_rel.drop_duplicates() #drop duplictes that arise from adding the mean column
+t11_rec = t11_rel_uniq.pivot('lambda','k','mean') #reshape to have a k by lambda table
+
+
+sns.set(font_scale=1.2)
+vacio = ["" for _ in xrange(4)]
[…]

### 2016-07-16 presentations/edindata/run-small.py

@@ -1,4 +1,6 @@
 import modsinglescalar as rmd
+import modsinglescalaronlyfitness as fit
+import modsinglescalaronlylearning as learner
 #parameter order for run(*) is run(alpha,cost,lam,k,sample_amount, learning_parameter,gens,runs)
 
 
@@ -18,7 +20,7 @@ def cost_to_learning():
 
 def ration_to_seq():
     a = 1
-    lam = [1,10,20,30,40,50]
+    lam = [x for x in xrange(1,50)]
     c = .4
     seq_length = [x for x in xrange(1,20)]
     sample_amount = 20
@@ -30,3 +32,36 @@ def ration_to_seq():
         for k in seq_length:
             rmd.run(a,c,l,k,sample_amount,learn,gens,runs)
 
+def alpha_to_seq():
+    a = [x/100. for x in xrange(1,101)]
+    lam = 20
+    c = .4
+    seq_length = [x for x in xrange(1,20)]
+    sample_amount = 20
+    learn = 3
+    gens = 20
+    runs = 1000
+    
+    for al in a:
+        for k in seq_length:
+            rmd.run(al,c,l,k,sample_amount,learn,gens,runs)
+
+def only_r_and_m():
+    a = 1
+    lam = 30
+    cost = [x/100. for x in xrange(100)]
+    learn = 1
+    seq_length = 5
+    sample_amount = 10
+    gens = 20
+    runs = 1000
+    
+    for c in cost:
+        print 'learn-only'
+        learner.run(a,c,lam,seq_length,sample_amount,learn,gens,runs)
+        print 'fit-only'
+        fit.run(a,c,lam,seq_length,sample_amount,learn,gens,runs)
+
+
+
+only_r_and_m()

### 2016-07-19 presentations/edindata/plots2.py
small presentation changes

@@ -109,6 +109,18 @@ plt.show()
 
 sys.exit()
 
+fig,axn = plt.subplots(1, 2, sharex=False, sharey=True)
+
+axn.flat[0] = sns.heatmap(t10_rec)
+axn.flat[1] = sns.heatmap(t11_rec)
+#for i, ax in enumerate(axn.flat):
+#    sns.heatmap(df, ax=ax,
+#                cbar=i == 0,
+#                vmin=0, vmax=1,
+#                cbar_ax=None if i else cbar_ax)
+
+fig.tight_layout(rect=[0, 0, .9, 1])
+plt.show()
 
 
 #### development over cost ###

### 2016-08-01 paper/paper.tex
paper introduction modifications

@@
 \tableofcontents
-\section{Introduction}\label{sec:introduction}
+\section{The semantics-pragmatics divide}\label{sec:introduction}
 
-\hl{some general sentence here}
 In linguistic theorizing, it is common to draw a distinction between semantics and pragmatics.
-Broadly speaking, semantics concerns the truth-conditional content of expressions, whereas pragmatics deals with information beyond such literal meanings.
+Broadly speaking, the former concerns the truth-conditional content of expressions, whereas the latter concerns information beyond such literal meanings.
 For instance, the integration of context and the derivation of defeasible inferences.
-Thus, the information conveyed is seldom if ever solely determined by semantics, but rather by its pragmatic enrichment.
-The present investigation addresses the selection of particular semantics over others in light of pragmatics.
-%  the semantics-pragmatics distinction raises the challenge of justifying the former's structure.
+Thus, the information conveyed by an utterance is seldom, if ever, solely determined by semantics, but rather by its pragmatic enrichment.
+For instance, truth-conditionally, {\em Bill read five books} does not set an upper-bound to the amount of books that Bill read; he may have read six, seven, or more books.
+However, the (defeasible) inference that {\em Bill read no more than five books} may be drawn on pragmatic grounds.
+After all, if Bill had read, for example, six books, and had the speaker known this, she would have said so.
+In this manner, what is conveyed (pragmatically) can go beyond what is (literally) said \citep{grice:1975}.
 
-Consider, for instance, the utterance {\em Bill read five books}.
-Truth-conditionally, this utterance does not set an upper-bound for the amount of books that Bill read; he may have read six, seven, or more books.
-Pragmatically, however, an addressee who assumes the speaker to be cooperative and knowledgeable may draw the (defeasible) inference that {\em Bill read no more than five books}.
-After all, had he read more, for example, six books, and the speaker had known this, she would have said {\em Bill read six books} instead.
-In this way, what is conveyed (pragmatically) goes beyond what is (literally) said \citep{grice:1975}.
+Much research in semantics and pragmatics is aimed at characterizing the information conveyed by classes of expressions such as numerals in terms of either domain, or their interplay.
+Thus, while there is disagreement on where their boundary lies, the distinction between semantics and pragmatics plays an important (if tacit) role in linguistics.
+However, an issue that has received little attention is the justification of semantic structure in light of pragmatics.
+Put differently, this issue concerns the selection and pervasiveness of certain semantic structures over others under consideration of the informational enrichment provided by pragmatics.
 
-Much research in semantics and pragmatics has been aimed at characterizing the information conveyed by classes of expressions, such as numerals, in terms of either domain, or their interplay.
-Thus, while there is disagreement on where their boundary lies, their distinction remains a core (if tacit) assumption in linguistics.
-An issue that has received little attention, however, is the justification of semantic structure in view of pragmatics.
-Prima facie, under a functionalist view, is is puzzling that {\em Bill read five books} does not semantically rule out its more informative or ``stronger'' alternatives.
-Wouldn't it serve speakers of a language better if weak(er) expressions such as {\em warm}, {\em or}, {\em some} or {\em big} were not truth-conditionally compatible with stronger alternatives such as {\em hot}, {\em and}, {\em all} and {\em huge}, respectively?
-After all, this is what an interlocutor's pragmatically enriched interpretation yields in the case of so-called scalar inferences.
+Prima facie, it is puzzling that {\em Bill read five books} does not semantically rule out its more informative or ``stronger'' alternatives.
+More pointedly, would it not serve language users better if weak(er) expressions such as {\em warm}, {\em or}, {\em some} or {\em big} were not truth-conditionally compatible with stronger alternatives such as, respectively, {\em hot}, {\em and}, {\em all} and {\em huge}?
+After all, analogously to the example sketched above, this is what a pragmatically enriched interpretation yields.
 Once a distinction between semantics and pragmatics is drawn, similar questions can be posed for other types of pragmatic inferences, such as ignorance inferences as (\ref{ex:ignorance}), or manner inferences such as (\ref{ex:manner-1}) and (\ref{ex:manner-2}).
-
-\begin{exe}
-\ex\label{ex:ignorance} \begin{xlist}
-	\ex Bill read at least five books
-	\ex $\leadsto$ Bill read five or more books (but the speaker doesn't know how many)
-\end{xlist}
-\ex \label{ex:manner-1} \begin{xlist}
-	\ex Mercader killed Trotsky.
-	\ex $\leadsto$ Mercader killed Trotsky in a stereotypical way
-    \end{xlist}
-\ex \label{ex:manner-2} \begin{xlist} \ex Mercader caused Trotsky to die
-		  \ex $\leadsto$ Mercader killed Trotsky in a non-stereotypical way
-    \end{xlist}
-\end{exe}
 

### 2016-08-02 paper/paper.tex
updated introduction draft for paper

@@
 After all, if Bill had read, for example, six books, and had the speaker known this, she would have said so.
-In this manner, what is conveyed (pragmatically) can go beyond what is (literally) said \citep{grice:1975}.
+In this manner, what is conveyed (pragmatically) can go beyond what is (literally) said.
 
@@
 Prima facie, it is puzzling that {\em Bill read five books} does not semantically rule out its more informative or ``stronger'' alternatives.
-More pointedly, would it not serve language users better if weak(er) expressions such as {\em warm}, {\em or}, {\em some} or {\em big} were not truth-conditionally compatible with stronger alternatives such as, respectively, {\em hot}, {\em and}, {\em all} and {\em huge}?
-After all, analogously to the example sketched above, this is what a pragmatically enriched interpretation yields.
-Once a distinction between semantics and pragmatics is drawn, similar questions can be posed for other types of pragmatic inferences, such as ignorance inferences as (\ref{ex:ignorance}), or manner inferences such as (\ref{ex:manner-1}) and (\ref{ex:manner-2}).
+More poignantly, would it not serve language users better if weak(er) expressions such as {\em warm}, {\em or}, {\em some} or {\em big} were truth-conditionally incompatible with stronger alternatives such as, respectively, {\em hot}, {\em and}, {\em all} and {\em huge}?
+After all, analogously to the example sketched above, this is what a pragmatically enriched interpretation yields; an upper-bound that rules out these alternatives through a so-called scalar inference \citep{horn:1972,gazdar:1979}.
+Once a distinction between semantics and pragmatics is drawn, similar questions can be posed for other types of pragmatic inferences, such as ignorance inferences as and manner inferences.
 
-In more general terms, the question then is what motivates the selection of particular semantic structures over other hypothetical alternatives in light of pragmatics.
-While a number of recent investigations have began to address issues pertaining to the development and selection of linguistic features, the semantics-pragmatics distinction has received little attention (see \citealt{steels:2015} and \citealt{tamariz+kirby:2016} for recent overviews).
-Our starting point is given by the overarching account of competing pressures that crystallizes across different frameworks: Natural languages need to allow for successful communication, one the one hand, and be well suited for acquisition on the other.
-Put differently, languages need to be well-adapted to the communicative needs within a linguistic community, and be learnable to survive their faithful transmission across generations.
+A number of recent investigations have began to address issues pertaining to the development and selection of linguistic properties (see \citealt{steels:2015} and \citealt{tamariz+kirby:2016} for recent overviews).
+Our starting point is given by the overarching account of competing pressures that has crystallized across these approaches: Natural languages need to allow for successful communication, one the one hand, and be well suited for acquisition, on the other.
+That is, they need to be well-adapted to the communicative needs within a linguistic community, as well as be learnable to survive their faithful transmission across generations.
+We proceed by instantiating these pressures in the replicator-mutator dynamics, commonly used in evolutionary game theory.
+Overall, this model allows for general and precise means to model the dynamics of linguistic pressures by combining functional pressure on successful communication, effects of learning biases on (iterated) Bayesian learning \citep{griffiths+kalish:2007}, and  probabilistic models of language use in populations with distinct lexica \citep{frank+goodman:2012,franke+jaeger:2014, bergen+etal:2016}.
+Drawing from the latter research strand allows for the representation of a distinction between semantics and pragmatics in communicative behavior.
+More broadly, it connects the recent surge in models of synchronic probabilistic rational language use with diachronic models of cultural evolution.
 
- \hl{Description of plan to lay out model and test it on upper-bounds.}
+\section{Conveying upper-bounds}
+\hl{start bottom up -- from the behavior of agents to the overall dynamics. Here, begin with language use and introduce scalar inferences on the way as the object of study we will focus on here}
 
-\hl{general notes on cultural evolution, biological evolution, and their connection}
+\paragraph{Language (fragments).}
+
+\paragraph{Signaling behavior.} With $\lambda \geq 1$ (rationality parameter), $\alpha \in [0,1]$ (pragmatic violations) and $pr \in \Delta(S)$ a common prior over $S$ (uniform so far):
+
+\begin{flalign}
+&R_{0}(s|m;L) \propto pr(s) L_{sm}\label{litl}\\
+&S_{0}(m|s;L) \propto \exp(\lambda \; L_{sm}) \label{lits}\\
+&R_{1}(s|m;L) \propto pr(s) S_{0}(m|s;L) \label{pragl}\\
+&S_{1}(m|s;L) \propto  \exp(\lambda \; R_{0}(s|m;L)^\alpha) \label{prags}
+\end{flalign}
 
@@
 In contrast, the iterated learning paradigm has focused on the effects of language transmission from generations of speakers to the next. \hl{explain IL}
-
-\subsection{Probabilistic models of language use}
-
-\paragraph{Language (fragments).}
-
-\paragraph{Signaling behavior.} With $\lambda \geq 1$ (rationality parameter), $\alpha \in [0,1]$ (pragmatic violations) and $pr \in \Delta(S)$ a common prior over $S$ (uniform so far):
-
-\begin{flalign}
-&R_{0}(s|m;L) \propto pr(s) L_{sm}\label{litl}\\
-&S_{0}(m|s;L) \propto \exp(\lambda \; L_{sm}) \label{lits}\\
-&R_{1}(s|m;L) \propto pr(s) S_{0}(m|s;L) \label{pragl}\\
-&S_{1}(m|s;L) \propto  \exp(\lambda \; R_{0}(s|m;L)^\alpha) \label{prags}
-\end{flalign}
 

### 2016-08-02 paper/paper.tex
paper section 2 draft update

@@
 \maketitle
-\tableofcontents
 \section{The semantics-pragmatics divide}\label{sec:introduction}
@@
 After all, analogously to the example sketched above, this is what a pragmatically enriched interpretation yields; an upper-bound that rules out these alternatives through a so-called scalar inference \citep{horn:1972,gazdar:1979}.
-Once a distinction between semantics and pragmatics is drawn, similar questions can be posed for other types of pragmatic inferences, such as ignorance inferences as and manner inferences.
+The present investigation focuses on the lack of lexicalization of such upper-bounds in natural languages.
+That is, we seek to provide an explanation for the selection of the particular semantics of scalar expressions, which in turn hinges on the semantics-pragmatics distinction for what is ultimately conveyed to interlocutors.
+However, once a distinction between semantics and pragmatics is drawn, similar questions can be posed for other types of pragmatic inferences, such as ignorance and manner inferences.
 
@@
 \section{Conveying upper-bounds}
+Scalar inferences involve the pragmatic derivation of an upper-bound for weak scalar expressions provided the hearer assumes the speaker to be knowledgeable and cooperative.
+A scale is an informativity induced ordering of relevant alternatives.
+Here, informativity is understood as entailment.
+For instance, {\em all students came} entails {\em most students came}, which in turn entails {\em some students came}.
+In this sense, {\em some} is weaker than {\em all}.
+The inference derived from utterances involving weaker alternatives is that stronger ones do not hold, e.g.
+{\em some students came} may be taken to convey that {\em not all students came}.
+
+\hl{say something about this pattern being widespread across languages. Give examples of classes of expressions in English, as alluded above (numerals, scalar adjectives, quantifiers)}
+
+Framed in Gricean terms \citep{grice:1975}, the hearer reasons about the speaker's choice of a weak alternative over a stronger one; had the speaker known that a stronger alternatives holds, she would have said so as this would have been more informative, since she did not, the hearer can infer that the stronger alternatives do not hold.
+Analogously, a speaker who reasons about her addressee may be rely on her to derive this inference.
+In this way, the strengthened, upper-bounded, state of affairs can be conveyed without codifiying the bound explicitly in the semantics.
+
+\paragraph{Function-based explanations.} Two ingredients are key to this derivation; (the assumptions of) cooperation and knowledge about the issue at hand.
+It is not difficult to imagine cases were one or both are not given.
+The speaker may (be assumed to) want to withhold information about the students' attendance, or may have left early without being able to verify the attendance to a satisfactory degree.
+
+\paragraph{Learning-based explanations.}
+
 \hl{start bottom up -- from the behavior of agents to the overall dynamics. Here, begin with language use and introduce scalar inferences on the way as the object of study we will focus on here}
@@
 
-\section{Upper-bounds in weak scalar alternatives}
-\hl{phenomenon description}
-
-\subsection{Analysis}
+\section{Analysis}
 

### 2016-08-04 paper/paper.tex
draft cogsci slides

--- 2016-08-02 paper/paper.tex
@@
 \section{Conveying upper-bounds}
-\hl{start bottom up -- from the behavior of agents to the overall dynamics. Here, begin with language use and introduce scalar inferences on the way as the object of study we will focus on here}
+Scalar inferences refer to the pragmatic derivation of an upper-bound for weak scalar expressions to the effect that stronger alternatives are inferred to not hold, e.g.
+{\em some students came} may be taken to convey that {\em not all students came}.
+The order of an expression such as {\em some} with respect to an alternative, e.g., {\em all}, is induced by entailment.
+For instance, {\em all students came} entails {\em most students came}, which in turn entails {\em some students came}.
+In this sense, {\em some} is weaker than {\em all}.
+A considerable class of natural language expressions do not lexicalize an upper-bound and can be ordered in this fashion, allowing for their pragmatic strengthening.
+As alluded at above, examples in English include numerals, scalar adjectives, quantifiers, modals, and connectives.
+\hl{possibly add some typological data on universality, frequency, monomorphemic status}
+
+The pragmatic enrichment of the semantic content of such expressions is enabled by mutual reasoning \citep{grice:1975}.
+More specifically, it is driven by interlocutors' mutual expectations of rational language use.
+The hearer reasons about the speaker's choice of a weak alternative over a stronger one.
+Had the speaker known that a stronger alternatives holds, she would have said so as this would have been more informative.
+Since she did not, the hearer can infer that the stronger alternative does not hold.
+Analogously, a speaker who reasons about her addressee may rely on her to derive this inference.
+In this way, a strengthened, upper-bounded, state of affairs can be conveyed without codifying the bound explicitly in the semantics.
+
+However, while pragmatics offers means to convey upper-bounds, the question why they are not part of the lexical meaning of these expressions remains.
+There are two main explanatory venues for this pattern.
+The first targets the functional advantages a lack of upper-bounds may confer to language users, whereas the second focuses on a learnability advantage of simpler over more complex semantic representations.
+
+\paragraph{Function-based explanations.} Two assumptions are key to the pragmatic strengthening of weak alternatives: (the assumption of) cooperation and knowledge about the issue at hand.
+That is, the hearer needs to assume the speaker to be as informative as possible, i.e., not to withhold information, and that the speaker is knowlegeable, e.g., she knows whether {\em all students came}.
+Conversely, the speaker needs to assume the hearer to regard these conditions as satisfied.
+It is not difficult to imagine scenarios where either or both of these conditions are not given.
+For instance, the speaker may (be assumed to) not want to disclose all information about the students' attendance, or may have left early without being able to verify the attendance to a satisfactory degree.
+\hl{here}
+
+\paragraph{Learning-based explanations.}
 
@@
     \item $\Phi = \sum_i x_i f_i$
-    \item $Q_{ij} \propto [\sum_d P(d|t_i) \; P(t_j|d)]^l$, where $P(t_j|d) \propto p^\star(t_j) P(d|t_j)$, $d$ is a sequence of observations of length $k$ of the form \tuple{\tuple{s_i,m_j}, ...
+    \item $Q_{ij} \propto \sum_d P(d|t_i) \; P(t_j|d)$, where $P(t_j|d) \propto [p(t_j) P(d|t_j)]^l$, $d$ is a sequence of observations of length $k$ of the form \tuple{\tuple{s_i,m_j}, ...
 \tuple{s_k, m_l}}, and $l \geq 1$ is a learning parameter.

### 2016-08-04 presentations/edindata/plot-fitness.py

@@ -1,0 +1,103 @@
+###
+import pandas as pd
+import numpy as np 
+import matplotlib.pyplot as plt
+import seaborn as sns
+#sns.set_context(rc={'lines.markeredgewidth': 0.5})
+import os
+import glob
+import sys
+###
+###Column headers:###
+#run_ID,t1_initial,t2_initial,t3_initial,t4_initial,t5_initial,t6_initial,t7_initial,t8_initial,t9_initial,t10_initial,t11_initial,t12_initial,
+#alpha,prior_cost_c,lambda,k,sample_amount,learning_parameter,generations,
+#t1_final,t2_final,t3_final,t4_final,t5_final,t6_final,t7_final,t8_final,t9_final,t10_final,t11_final,t12_final
+#####################
+
+
+print 'Loading data fitness'
+path =r'./results-fit' 
+all_files = glob.glob(os.path.join(path, "*.csv"))     # advisable to use os.path.join as this makes concatenation OS independent
+df = pd.concat(pd.read_csv(f) for f in all_files)
+
+print 'Loading data learn'
+path =r'./results-learn' 
+all_files = glob.glob(os.path.join(path, "*.csv"))     # advisable to use os.path.join as this makes concatenation OS independent
+dl = pd.concat(pd.read_csv(f) for f in all_files)
+
+def cust_mean(grp):
+    grp['mean'] = grp['t11_final'].mean()
+    return grp
+
+
+
+
+### development over cost ###
+
+a = 1
+k = 5
+lam = 30
+sample = 10
+learn = 1
+
+
+final_groupA = df.loc[:,'alpha':]
+final_groupB = dl.loc[:,'alpha':]
+
+groupA = final_groupA.loc[final_groupA['alpha'] == a]
+groupA = groupA.loc[groupA['k'] == k]
+groupA = groupA.loc[groupA['sample_amount'] == sample]
+groupA = groupA.loc[groupA['learning_parameter'] == learn]
+groupA = groupA.loc[groupA['prior_cost_c'] == 0.1]
+
+t_finalA = groupA.groupby(['lambda'])
[…]

### 2016-08-04 run-small.py
minor changes simulation script

@@ -1,17 +1,15 @@
 import modsinglescalar as rmd
-#parameter order for run(*) is run(alpha,cost,lam,k,sample_amount, learning_parameter,gens,runs)
+#parameter order for run(*) is run(alpha,cost,lam,k, learning_parameter,gens,runs)
 
-a = 1
-lam = [10,30]
-cost = [0.01,0.05,0.1,0.2,0.3,0.4,0.5,0.7,0.8,0.9]
+alpha = 1
+lam = 30
+cost = [x/100. for x in xrange(1,101)]
 seq_length = [1,3,5,9,13,15,17,20]
 learning_parameter = [1,2,4,6,8,10]
 gens = 20
-runs = 1000
+runs = 10000
 
 for c in cost:
-    for l in lam:
-        for k in seq_length:
-            for sampl in sample_amount:
-                for learn in learning_parameter:
-                    rmd.run(a,c,l,k,learn,gens,runs)
+    for k in seq_length:
+        for learn in learning_parameter:
+                    rmd.run(alpha,c,lam,k,learn,gens,runs)
